---
title: "SAE_analysis 3rd version"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

# Set up ----

```{r setup, message=FALSE}
rm(list=ls())

path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
library(ggplot2)
library(sf)
library(viridis)
library(scales)
library(patchwork)
library(rlang)
library(gt)
```

# Import datasets ----

```{r cars, message=FALSE}
# SAE data
SAE_2024 <- read_csv2("../data/processed/SAE_final_2024.csv")

# ID data
id_raw <- read_csv2("../data/raw/SAE_2024/ID_2024r.csv")

# Geographical data
communes <- st_read("../data/raw/communes-1000m.geojson")
departements_map <- st_read("../data/raw/departements.geojson")

# Enquete Urgences data
enquete_urgences <- read_csv2("../data/raw/enquete_urgences.csv")
```

```{r}
# Select useful variables in id
id_raw <- id_raw %>%
  dplyr::select(fi, COMINSEE)

# Merge with the main dataset
SAE_2024 <- SAE_2024 %>%
  left_join(id_raw, by = c("finess" = "fi"))

# Select useful variables in communes
communes <- communes %>%
  dplyr::select(code, geometry)

# Merge with the geographical data
SAE_2024 <- SAE_2024 %>%
  left_join(communes, by = c("COMINSEE" = "code"))

# Count percentage of "MULTIPOLYGON EMPTY" in geometry
empty_geom_percentage <- sum(st_is_empty(SAE_2024$geometry)) / nrow(SAE_2024) * 100

# Select useful variables in enquete urgences
enquete_urgences <- enquete_urgences %>%
  dplyr::select(FINESS, DENSITE_GENERALISTES_LIB)

# Merge with the main dataset
SAE_2024 <- SAE_2024 %>%
  left_join(enquete_urgences, by = c("finess" = "FINESS"))
```

# Descriptive statistics ----

```{r}
# Add the number of liberal physicians to the etp
SAE_2024 <- SAE_2024 %>%
  mutate(
    # Convert liberals number to ETP
    etp_liberals = number_physicians_liberal,
    # Sum the ETP of employed physicians and liberals
    etp_total_combine = etp_physicians_empl_emergency + etp_liberals,
  )

# Hospital 570026682 seems to be an outlier, replace etp_total_combine by the median
SAE_2024 <- SAE_2024 %>%
  mutate(etp_total_combine = ifelse(finess == "570026682", median(etp_total_combine, na.rm = TRUE), etp_total_combine))

# Count percentage of NA
na_percentage <- sum(is.na(SAE_2024$etp_total_combine)) / nrow(SAE_2024) * 100
na_percentage_empl <- sum(is.na(SAE_2024$etp_physicians_empl_emergency)) / nrow(SAE_2024) * 100
```



```{r}
# Plot the density function of etp_total_combine (add median vertical line)
ggplot(SAE_2024, aes(x = etp_total_combine)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  geom_vline(xintercept = median(SAE_2024$etp_total_combine, na.rm = TRUE), linetype = "dashed", color = "red") +
  labs(title = "Density of the Number of Doctors in Emergency Services",
       x = "Number of Doctors (ETP)",
       y = "Density") +
  theme_minimal()

# Median ETP
median_etp <- median(SAE_2024$etp_total_combine, na.rm = TRUE)

# Export the plot
ggsave("../results/Final_pres/doctors_density.png", width = 10, height = 8)
```

```{r}
# Compute density to find the mode (value with the highest density)
density_data <- density(SAE_2024$etp_total_combine, na.rm = TRUE)
mode_etp <- density_data$x[which.max(density_data$y)]
median_etp <- median(SAE_2024$etp_total_combine, na.rm = TRUE)

# Graph with both median and mode
ggplot(SAE_2024, aes(x = etp_total_combine)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  # Median line
  geom_vline(xintercept = median(SAE_2024$etp_total_combine, na.rm = TRUE), 
             linetype = "dashed", color = "red") +
  # Mode line
  geom_vline(xintercept = mode_etp, 
             linetype = "dashed", color = "darkgreen") +
  labs(title = "Density of the Number of Doctors in Emergency Services",
       x = "Number of Doctors (ETP)",
       y = "Density") +
  theme_minimal()

# Export the plot
ggsave("../results/Final_pres/doctors_density.png", width = 10, height = 8)

# Compute the percentage of the data that's below the mode
percentage_below_mode <- sum(SAE_2024$etp_total_combine < mode_etp, na.rm = TRUE) / nrow(SAE_2024) * 100
```

## Point map

We want to have an idea of whether the hospitals with few doctors are concentrated in specific areas of France, or if they are spread out across the country.

```{r}
# Convert SAE_2024 to sf object
hospitals_clean <- SAE_2024 %>%
  st_as_sf() %>% 
  st_transform(2154) %>% # Project to meters
  filter(!st_is_empty(geometry) & !is.na(etp_total_combine)) %>% 
  st_make_valid() %>% 
  st_centroid()

# Plot (remove the coordinates axes and gridlines for a cleaner look)
etp_map <-  ggplot() +
  geom_sf(data = st_transform(departements_map, 2154), fill = "white", color = "gray60") +
  geom_sf(data = hospitals_clean, aes(color = etp_total_combine), size = 2, shape = 3) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Number of Doctors (ETP)") +
  coord_sf(datum = NA) + # Use the projected CRS for the map view
  theme_minimal() +
  theme(
    # Suppress axes
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    # Suppress gridlines
    panel.grid = element_blank(),
    # Suppress panel border
    panel.border = element_blank()
  ) +
  # Add title (centered)
  labs(title = "Geographical Distribution of Emergency Services in France")

# Export
ggsave("../results/Final_pres/doctors_point_map.png", etp_map, width = 10, height = 8)
```

Zoom-in on Ile-de-France for better readability.

```{r}
# Zoom in on ile de france (region_number = 11)
hospitals_idf <-  hospitals_clean %>%
  filter(region_number == 11)

idf_map <- departements_map %>%
  filter(substr(code, 1, 2) %in% c("75", "77", "78", "91", "92", "93", "94", "95"))

# Plot
etp_map_idf <-  ggplot() +
  geom_sf(data = st_transform(idf_map, 2154), fill = "white", color = "gray60") +
  geom_sf(data = hospitals_idf, aes(color = etp_total_combine), size = 2, shape = 3) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Number of Doctors (ETP)") +
  coord_sf(datum = NA) + 
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank()
  ) +
  labs(title = "Geographical Distribution of Emergency Services in Île-de-France")

# Export
ggsave("../results/Final_pres/doctors_point_map_idf.png", etp_map_idf, width = 10, height = 8)
```


## Correlation between number of doctors and other dimensions

Juridical status

```{r}
# Boxplot of number of doctors by juridical status
ggplot(SAE_2024, aes(x = juridical_status, y = etp_total_combine, fill = juridical_status)) +
  geom_boxplot(fill = "steelblue") +
  labs(title = "Distribution of the Number of Doctors by Juridical Status",
       x = "Juridical Status",
       y = "Number of Doctors (ETP)") +
  theme_minimal() +
  theme(legend.position = "none")

# Export
ggsave("../results/Final_pres/doctors_js.png", width = 10, height = 8)

# Anova test to check if the differences in the number of doctors across juridical status are statistically significant
anova_result <- aov(etp_physicians_empl_emergency ~ juridical_status, data = SAE_2024)
```

Type

```{r}
# Boxplot of number of doctors by type of emergency service (add labels to the x axis)
SAE_2024 <- SAE_2024 %>%
  mutate(emergency_type_label = case_when(
    emergency_type == 1 ~ "Pediatric",
    emergency_type == 2 ~ "General",
    emergency_type == 3 ~ "Both",
    TRUE ~ "NA"
  ))

ggplot(SAE_2024, aes(x = emergency_type_label, y = etp_total_combine)) +
  geom_boxplot(fill = "steelblue") +
  labs(title = "Distribution of the Number of Doctors by Type of Emergency Service",
       x = "Type of Emergency Service",
       y = "Number of Doctors (ETP)") +
  theme_minimal() +
  theme(legend.position = "none")

# Export
ggsave("../results/Final_pres/doctors_type.png", width = 10, height = 8)

# Anova test to check if the differences in the number of doctors across types of emergency service are statistically significant
anova_result_type <- aov(etp_physicians_empl_emergency ~ as.factor(emergency_type), data = SAE_2024)
```
Size of the hospital in terms of number of visits

```{r}
# Scatter plot of number of doctors by number of visits
ggplot(SAE_2024, aes(x = total_emergency_visits_win, y = etp_total_combine)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  scale_x_continuous(labels = comma) +
  labs(title = "Correlation between the Number of Doctors and the Number of Visits",
       x = "Number of Visits in Emergency Services",
       y = "Number of Doctors (ETP)") +
  theme_minimal()

# Export
ggsave("../results/Final_pres/doctors_visits.png", width = 10, height = 8)


# Correlation test to check if the correlation between the number of doctors and the number of visits is statistically significant
cor_test <- cor.test(SAE_2024$total_emergency_visits_win, SAE_2024$etp_total_combine, method = "pearson")
```

Densité de médecins généralistes libéraux

```{r}
# Scatter plot of number of doctors by density of liberals
ggplot(SAE_2024, aes(x = as.numeric(DENSITE_GENERALISTES_LIB), y = etp_total_combine)) +
  geom_point(alpha = 0.6) +
  labs(title = "Correlation between the Number of Doctors and the Density of Liberal General Practitioners",
       x = "Density of Liberal General Practitionners in the Department",
       y = "Number of Doctors (ETP)") +
  theme_minimal()

# Correlation test to check if the correlation between the number of doctors and the density
cor_test_lib <- cor.test(as.numeric(SAE_2024$DENSITE_GENERALISTES_LIB), SAE_2024$etp_total_combine, method = "pearson")
cor_test_lib

```
NOT SIGNIF. DON'T INCLUDE

The 3 first results are all stat signif, highlighting the necesity to control for them in the regression.

```{r}
# 1. Define helper to extract data from ANOVA and Correlation
anova_res_status <- summary(anova_result)[[1]]
anova_res_type <- summary(anova_result_type)[[1]]

# 2. Build the data frame
results_table <- data.frame(
  Variable = c("Juridical Status", "Emergency Type", "Number of Visits"),
  Test_Type = c("ANOVA", "ANOVA", "Correlation"),
  F_Stat = c(round(anova_res_status$`F value`[1], 2), round(anova_res_type$`F value`[1], 2), NA),
  Pearson_r = c(NA, NA, round(cor_test$estimate, 3)),
  P_Value = c(anova_res_status$`Pr(>F)`[1], anova_res_type$`Pr(>F)`[1], cor_test$p.value)
)

# 3. Create the gt table
results_table %>%
  gt() %>%
  tab_header(
    title = "Statistical Analysis Summary",
    subtitle = "Drivers of ED Physician Staffing"
  ) %>%
  cols_label(
    Variable = "Dimension",
    Test_Type = "Test",
    F_Stat = "F-Statistic",
    Pearson_r = "Pearson r",
    P_Value = "p-value"
  ) %>%
  # Replace NAs with a dash for cleaner presentation
  sub_missing(columns = everything(), missing_text = "-") %>%
  fmt_number(columns = c(F_Stat, Pearson_r, P_Value), decimals = 4) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  # Highlight significant results
  tab_style(
    style = cell_fill(color = "white"),
    locations = cells_body(rows = P_Value < 0.05)
  )

# Export the table as a PNG image
results_table %>%
  gt() %>%
  tab_header(
    title = "Statistical Analysis Summary",
    subtitle = "Drivers of ED Physician Staffing"
  ) %>%
  cols_label(
    Variable = "Dimension",
    Test_Type = "Test",
    F_Stat = "F-Statistic",
    Pearson_r = "Pearson r",
    P_Value = "p-value"
  ) %>%
  # Replace NAs with a dash for cleaner presentation
  sub_missing(columns = everything(), missing_text = "-") %>%
  fmt_number(columns = c(F_Stat, Pearson_r, P_Value), decimals = 4) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  # Highlight significant results
  tab_style(
    style = cell_fill(color = "white"),
    locations = cells_body(rows = P_Value < 0.05)
  ) %>%
  gtsave("../results/Final_pres/stat_analysis_summary.png")
```
