---
title: "SAE_cleaning_aggregation"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

The last data cleaning thing we have to do is aggregate the hospitals that have both pediatric and general emergencies into one row.


```{r setup, include=FALSE}
rm(list=ls())

path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
library(ggplot2)
library(sf)
library(viridis)
library(scales)
library(tibble)
```

# Import datasets ----

```{r cars, message=FALSE}
SAE_2024 <- read_csv2("../data/intermediate/SAE_cleaned_outliers_2024.csv")
Legend <- read_csv2("../docs/SAE_2024/Legends_short.csv")
```

# Aggregate pediatric and general emergencies ----

```{r}
# Identify establishments with duplicate names
duplicate_names <- SAE_2024 %>%
  group_by(establishment_name) %>%
  filter(n() > 1) %>%
  pull(establishment_name) %>%
  unique()

# Create a dataset of the duplicate establishments for inspection
inspection_duplicates <- SAE_2024 %>%
  filter(establishment_name %in% duplicate_names) %>%
  arrange(establishment_name)
```

- Information variables (year	finess	establishment_name	juridical_status	region_number	region_name	dep_number	dep_name	town_name	postal_code): keep the first

- Capacity: drop the values, we will compute it again after the aggregation.

- Sum total_hoptialized_emergency, beds_number, total_emergency_visits_raw.

- All the following variables are the same for the duplicates. Compute the mean (in case we overlooked a difference) hours_physician_total_raw, hours_physicians_emergency_raw	hours_nurses_total_raw	hours_nurses_emergency_raw	hours_paramedics_raw	hours_regulation_physicians, hours_regulation_assistants,	etp_physicians_empl_total	etp_physicians_empl_emergency,	number_physicians_liberal	number_medecine_interns,	number_medecine_junior_doctors,	number_emergency_interns	number_emergency_junior_doctors, hours_physicians_total_mod	hours_physicians_emergency_mod	hours_nurses_total_mod	hours_nurses_emergency_mod	hours_regulation_assistants_raw	hours_regulation_assistants_mod	hours_paramedics_mod	etp_total_win	etp_emergency_win.

- Dummies: keep the highest: waiting_monitoring, joint_personnel_SMUR_emergency, dummy_regulation_physicians	dummy_paramedics	dummy_regulation_assistants	dummy_number_liberal_phyisicians	dummy_joint_personnel, dummy_medecine_interns	dummy_emergency_interns	dummy_medecine_junior_doctors	dummy_emergency_junior_doctors.

- Rates (hosp_rate_raw,	hosp_rate_mod): erase values. We will compute them again after the aggregation.

- Winsorized/modified variables: drop total_emergency_visits_win and hosp_rate_mod. Apply the same correction after aggregation.

We also replace the values of the emergency_type variable:
- 1 if only PED,
- 2 if only GEN,
- 3 if both PED and GEN.

```{r}
SAE_final <- SAE_2024 %>%
  group_by(establishment_name) %>%
  summarise(
    # Information variables
    across(c(year, finess, juridical_status, region_number, region_name, 
             dep_number, dep_name, town_name, postal_code), first),
    
    # Flow and beds variables
    total_hospitalized_emergency = sum(total_hospitalized_emergency, na.rm = TRUE),
    beds_number = sum(beds_number, na.rm = TRUE),
    total_emergency_visits_raw = sum(total_emergency_visits_raw, na.rm = TRUE),
    
    # Personnel and hours: mean
    across(c(
      hours_physicians_total_raw, hours_physicians_emergency_raw, hours_nurses_total_raw, 
      hours_nurses_emergency_raw, hours_paramedics_raw, hours_regulation_physicians, 
      hours_regulation_assistants, etp_physicians_empl_total, etp_physicians_empl_emergency, 
      number_physicians_liberal, number_medecine_interns, number_medecine_junior_doctors, 
      number_emergency_interns, number_emergency_junior_doctors, hours_physicians_total_mod, 
      hours_physicians_emergency_mod, hours_nurses_total_mod, hours_nurses_emergency_mod, 
      hours_regulation_assistants_raw, hours_regulation_assistants_mod, hours_paramedics_mod, 
      etp_total_win, etp_emergency_win
    ), function(x) mean(x, na.rm = TRUE)),
    
    # Dummies & Regulation : Safe Max to avoid -Inf
    across(c(
      regulation, waiting_monitoring, joint_personnel_SMUR_emergency, dummy_regulation_physicians, 
      dummy_paramedics, dummy_regulation_assistants, dummy_number_liberal_phyisicians, 
      dummy_joint_personnel, dummy_medecine_interns, dummy_emergency_interns, 
      dummy_medecine_junior_doctors, dummy_emergency_junior_doctors
    ), function(x) {
      if(all(is.na(x))) return(NA_real_)
      max(x, na.rm = TRUE)
    }),
    
    # Emergency type
    emergency_type = case_when(
      any(emergency_type == "GEN") & any(emergency_type == "PED") ~ 3,
      all(emergency_type == "PED") ~ 1,
      all(emergency_type == "GEN") ~ 2,
      TRUE ~ 2
    ),
    
    .groups = "drop"
  ) %>%
  
  # New winsorization based on aggregated raw visits
  mutate(
    total_emergency_visits_win = ifelse(
      total_emergency_visits_raw > quantile(total_emergency_visits_raw, 0.99, na.rm = TRUE),
      quantile(total_emergency_visits_raw, 0.99, na.rm = TRUE),
      total_emergency_visits_raw
    )
  ) %>%
  
  # Dual Rate computation
  mutate(
    hosp_rate_raw = (total_hospitalized_emergency / total_emergency_visits_raw) * 100,
    hosp_rate_mod = (total_hospitalized_emergency / total_emergency_visits_win) * 100
  )
```
```{r}
# Check that the number of hospitals in SAE_final is the same as the number of unique establishment names in SAE_2024
length(unique(SAE_2024$establishment_name)) == nrow(SAE_final)
```

We have the same number of variables as in SAE_2024 minus two, which is coherent.

We still need to recalculate the capacity variable.

```{r}
# Compute new terciles of the beds_number variable in SAE_final
terciles_final <- quantile(SAE_final$beds_number, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)

# Create new capacity variable based on the new terciles
SAE_final <- SAE_final %>%
  mutate(capacity = case_when(
    beds_number <= terciles_final[2] ~ "Low capacity",
    beds_number <= terciles_final[3] ~ "Medium capacity",
    beds_number >  terciles_final[3] ~ "High capacity"
  )) %>%
  mutate(capacity = factor(capacity, 
                           levels = c("Low capacity", "Medium capacity", "High capacity")))
```

We still need to apply outlier correction to hosp_rate.

```{r}
# Create outlier detection function
find_outliers <- function(data, variable) {
  # Compute stats
  stats <- data %>%
    summarise(
      q1 = quantile({{variable}}, 0.25, na.rm = TRUE),
      q3 = quantile({{variable}}, 0.75, na.rm = TRUE),
      iqr = q3 - q1
    )
  upper_limit <- stats$q3 + 1.5 * stats$iqr
  lower_limit <- stats$q1 - 1.5 * stats$iqr
  # Return results
  data %>%
    filter({{variable}} > upper_limit | {{variable}} < lower_limit) %>%
    select(establishment_name, beds_number, capacity, emergency_type, juridical_status, {{variable}})
}

# Apply to hosp_rate_raw
outliers_hosp_rate <- find_outliers(SAE_final, hosp_rate_raw)
```

Following the exact same logic as in the outlier cleaning code, we remove the two same outliers from hosp_rate.

```{r}
# Replace outliers with median value for CH CHAUNY and CH DE VIERZON
SAE_final <- SAE_final %>%
  mutate(
    hosp_rate_mod = ifelse(
      establishment_name %in% c("CH CHAUNY", "CH DE VIERZON"),
      median(hosp_rate_raw, na.rm = TRUE),
      hosp_rate_raw
    )
  )
```

We now have the exact same number of variables as in the SAE_2024 dataset.

We reorder the variables for clarity.

```{r}
# Reordering variables for readability
SAE_final <- SAE_final %>%
  select(
    # 1. Identification & Geography
    year, establishment_name, finess, 
    region_number, region_name, dep_number, dep_name, town_name, postal_code,
    
    # 2. Hospital Structure & Characteristics
    juridical_status, capacity, beds_number, emergency_type, regulation,
    
    # 3. Activity Metrics (Raw & Processed)
    total_emergency_visits_raw, total_emergency_visits_win,
    total_hospitalized_emergency, 
    hosp_rate_raw, hosp_rate_mod,
    
    # 4. Staffing - Physicians & Junior Doctors
    number_physicians_liberal, 
    number_medecine_interns, number_medecine_junior_doctors,
    number_emergency_interns, number_emergency_junior_doctors,
    etp_physicians_empl_total, etp_physicians_empl_emergency,
    
    # 5. Hours & ETP (Raw & Modified)
    starts_with("hours_physicians"),
    starts_with("hours_nurses"),
    starts_with("hours_paramedics"),
    starts_with("hours_regulation"),
    starts_with("etp_"),
    
    # 6. Dummies (Binary indicators)
    starts_with("dummy_"),
    waiting_monitoring, joint_personnel_SMUR_emergency,
    
    # Everything else (if any were missed)
    everything()
  )
```

```{r}
# Save the final cleaned and aggregated dataset
write_csv2(SAE_final, "../data/processed/SAE_final_2024.csv")
```