---
title: "SAE_cleaning"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

# Set up ----

```{r setup, include=FALSE}
path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(jsonlite)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
```

# Import 2024 datasets ----

```{r cars, message=FALSE}
identification1 <- read_delim("../data/raw/SAE_2024/ID_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

identification2 <- read_delim("../data/raw/SAE_2024/IDB_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

urgences1 <- read_delim("../data/raw/SAE_2024/URGENCES_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

urgences2 <- read_delim("../data/raw/SAE_2024/URGENCES2_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

personnel_urg1 <- read_delim("../data/raw/SAE_2024/PCAMEDURG_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

personnel_urg2 <- read_delim("../data/raw/SAE_2024/PCAMEDURG_P_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

personnel_all <- read_delim("../data/raw/SAE_2024/Q20_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

personnel_interns <- read_delim("../data/raw/SAE_2024/Q22_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))

synthesis <- read_delim("../data/raw/SAE_2024/SYGEN_2024r.csv", delim = ";", locale = locale(decimal_mark = "."))
```


# Clean and merge all 2024 tables ----

## identification1 & identification2

```{r, message=FALSE}
# Remove unnecessary columns
identification1 <- identification1 %>%
  select(-c("bor", "rscom", "grp", "espic", "ccr", "nat", "sir", "reg_diff", "dep_diff", "ARMT", "ETAT_SAISIE", "VAGUE", "PROVENANCE", "NUMVOI", "TYPVOI", "NOMVOI", "BP", "COMPD", "FILIALE0", "SIRTET", "RSTET", "MODEJ", "FINRAT"), -starts_with(c("FIAUT", "RSAUT", "DADS_", "TYPMD", "FIFU", "SCFU", "VAL", "RESPREG")))

identification2 <- identification2 %>%
  select(-c("BOR", "GRP", "ESPIC", "CCR", "NAT", "SIR", "ARMT", "VAGUE", "NUMTEL", "COMPRS", "NUMVOI", "TYPVOI", "ADRVOIENOM", "COMPDist", "ADRLIBCOM", "COMINSEE", "ADRBP"))

# Standardize column names
identification1 <- janitor::clean_names(identification1)
identification2 <- janitor::clean_names(identification2)

# Harmonize column names and add a source column
identification2 <- identification2 %>%
  rename(cpo = adrcpost) %>%
  mutate(source = "id2")
identification1 <- identification1 %>%
  mutate(source = "id1")

# Convert stj to a character in identification2
identification2 <- identification2 %>%
  mutate(stj = as.character(stj))

# Combine both identification tables
identification <- bind_rows(identification1, identification2)
```

## urgences1

```{r, message=FALSE}
# Remove unnecessary columns
urgences1 <- urgences1 %>%
  select(-"BOR")

# Standardize column names
urgences1 <- janitor::clean_names(urgences1)

# Check that all the FINESS numbers in urgences1 are in identification
all(urgences1$fi %in% identification$fi)

# Check that no table has duplicates
n_distinct(urgences1$fi, na.rm = TRUE) ==
  sum(!is.na(urgences1$fi))
n_distinct(identification$fi, na.rm = TRUE) ==
  sum(!is.na(identification$fi))

# Check NAs
sum(is.na(urgences1$fi))
sum(is.na(identification$fi))

# Print columns that are in both tables
intersect(names(urgences1), names(identification))

urgences1 <- urgences1 %>%
  janitor::clean_names() %>%
  # Pivot longer to have one row per type of emergency activity (general or pediatric) and per medical structure
  pivot_longer(
    cols = c(autgen, autped), 
    names_to = "urg_type", 
    values_to = "passu_check"
  ) %>%
  # Harmonize the "urg" variable to have the same values for both types of emergency activity (general or pediatric)
  mutate(urg = case_when(
    urg_type == "autgen" ~ "GEN",
    urg_type == "autped" ~ "PED",
    TRUE ~ NA_character_
  )) %>%
  # Keep only the rows where passu_check is not NA and > 0
  filter(!is.na(passu_check) & passu_check > 0)

# Merge and get rid of possible duplicates
SAE_main <- identification %>%
  inner_join(
    urgences1 %>%
      select(-any_of(setdiff(intersect(names(urgences1), names(identification)), "fi"))),
    by = "fi"
  )

# Check that some rows have id2 source
SAE_main %>%
  filter(source == "id2") %>%
  nrow() # 0

# Remove "source" column
SAE_main <- SAE_main %>%
  select(-"source")
```

## urgences2

```{r, message=FALSE}
# Remove unnecessary columns
urgences2 <- urgences2 %>%
  select(-c("BOR", "CONVURG24", "FICONV", "IMPL_SMUR_SITE"), -starts_with(c("OUV_", "AMPL_")))

# Standardize column names
urgences2 <- janitor::clean_names(urgences2)

# Check that all the FINESS numbers are in both tables
all(SAE_main$fi %in% urgences2$fi)

# Print columns that are in both tables
intersect(names(SAE_main), names(urgences2))

# Rename column that has the same name but a different meaning
urgences2 <- urgences2 %>%
  rename(regulation = reg)

# Merge and get rid of possible duplicates
SAE_main <- SAE_main %>%
  inner_join(
    urgences2 %>%
      select(-any_of(setdiff(intersect(names(urgences2), names(SAE_main)), c("fi", "urg")))),
    by = c("fi", "urg")
  )
```

NB: The following variables were dropped because of a too small number of observations (even if they were interesting):
IMPL_SMUR_SITE	Disposez-vous d’une implantation SMUR sur le même site ou avez-vous une convention/organisation formalisée ?
CONVURG24	Etes-vous conventionné avec une structure des urgences H24 ?
FICONV	Si oui, FINESS géographique de la structure des urgences (autorisée H24) signataire de la convention
OUV_SEMAINE	Êtes-vous ouvert du lundi au vendredi ?
AMPL_SEMAINE	Quelle est votre amplitude horaire journalière moyenne du lundi au vendredi ?
OUV_SAM	Êtes-vous ouvert le samedi ?
AMPL_SAM	Quelle est votre amplitude horaire moyenne le samedi ?
OUV_DIMFERIE	Êtes-vous ouvert le dimanche et les jours fériés ?
AMPL_DIMFERIE	Quelle est votre amplitude horaire journalière moyenne le dimanche et les jours fériés ?

## personnel_urg1

```{r, message=FALSE}
# Remove unnecessary columns
personnel_urg1 <- personnel_urg1 %>%
  select(-c("BOR"))

# Standardize column names
personnel_urg1 <- janitor::clean_names(personnel_urg1)

# Print columns that are in both tables
intersect(names(personnel_urg1), names(SAE_main))

# Merge and get rid of possible duplicates
SAE_main <- SAE_main %>%
  left_join( # keep the medical structure even if it has no personnel_urg1 data
    personnel_urg1 %>%
      select(-any_of(setdiff(intersect(names(SAE_main), names(personnel_urg1)), "fi"))),
    by = "fi"
  )
```

## personnel_urg2

```{r}
# Remove unnecessary columns
personnel_urg2 <- personnel_urg2 %>%
  select(-"BOR")

# Keep only the rows where PERSO = "M9999" (total personnel medical)
personnel_urg2 <- personnel_urg2 %>%
  filter(PERSO == "M9999")

# Remove "PERSO" and "MOD" columns
personnel_urg2 <- personnel_urg2 %>%
  select(-c("PERSO", "MOD"))

# Standardize column names
personnel_urg2 <- janitor::clean_names(personnel_urg2)

# Print columns that are in both tables
intersect(names(personnel_urg2), names(SAE_main))

# Merge and get rid of possible duplicates
SAE_main <- SAE_main %>%
  left_join( # keep the medical structure even if it has no personnel_urg2 data
    personnel_urg2 %>%
      select(-any_of(setdiff(intersect(names(SAE_main), names(personnel_urg2)), "fi"))),
    by = "fi"
  )
```

## personnel_interns

```{r}
personnel_interns <- personnel_interns %>%
  select(-"BOR")

# Standardize column names
personnel_interns <- janitor::clean_names(personnel_interns)

# Print columns that are in both tables
intersect(names(personnel_interns), names(SAE_main))

# Merge and get rid of possible duplicates
SAE_main <- SAE_main %>%
  left_join( # keep the medical structure even if it has no interns
    personnel_interns %>%
      select(-any_of(setdiff(intersect(names(SAE_main), names(personnel_interns)), "fi"))),
    by = "fi"
  )
```

## personnel_all

```{r, message=FALSE}
# Standardize column names
personnel_all <- janitor::clean_names(personnel_all)

# Keep only useful columns
personnel_all <- personnel_all %>%
  select(c("fi", "perso", "effsalplh", "effsalplf", "effsalpah", "effsalpaf", "efflibplh", "efflibplf", "efflibpah", "efflibpaf", "etpsalh", "etpsalf", "effsal", "etpsal", "efflib"))

# Keep only the observations if they correspond to a FINESS for which perso = "M1340" (médecine d'urgence)
personnel_all <- personnel_all %>%
  group_by(fi) %>%
  filter(any(perso == "M1340")) %>%
  ungroup() %>%
  filter(perso %in% c("M1340", "M9999"))

# Create new variables from the information in observations where perso = "M9999" (total personnel medical)
personnel_all <-  personnel_all %>%
  pivot_wider(
    names_from = perso,
    values_from = where(is.numeric)
  ) %>%
  rename_with(~ paste0("tot_", gsub("_M9999", "", .x)), ends_with("_M9999")) %>%
  rename_with(~ gsub("_M1340", "", .x), ends_with("_M1340")) %>%
  group_by(fi) %>%
  summarise(across(everything(), ~ first(na.omit(.x))), .groups = "drop")
```

Legend of the remaining columns:
- EFFSALPLH = effectif des personnels salariés temps plein hommes
- EFFSALPLF = effectif des personnels salariés temps plein femmes
- EFFSALPAH = effectif des personnels salariés stemps partiel hommes
- EFFSALPAF = effectif des personnels salariés stemps partiel femmes
- EFFLIBPLH = effectif des praticiens libéraux temps plein hommes
- EFFLIBPLF = effectif des praticiens libéraux temps plein femmes
- EFFLIBPAH = effectif des praticiens libéraux temps partiel hommes
- EFFLIBPAF = effectif des praticiens libéraux temps partiel femmes
- ETPSALH = ETP moyen annuel total des personnels salariés hommes
- ETPSALF = ETP moyen annuel total des personnels salariés femmes
- EFFSAL = effectif des personnels salariés total (somme des quatre EFFSAL)
- ETPSAL = somme de ETPSALH et ETPSALF
- EFFLIB = effectif des praticiens libéraux total (somme de tous les EFFLIB)

We choose not to include personnel_all in the final dataset since it does not represent the effective medical staff dedicated to the emergency services, contrarly to personnel_urg1 and personnel_urg2.

## synthesis

```{r}
columns_to_keep = c("FI", "FI_EJ", "AN", "ANT", "AUTGEN", "AUTPED", "EFFSAL_TOT", "EFFSAL_DT_URG", "EFFLIB_TOT", "EFFLIB_DT_URG", "PASSU_GEN", "PASSU_PED", "EFFPNM_TOT_PNM", "ETPSAL_DT_URG", "ETPSAL_TOT", "ETPPNM_TOT_PNM")

synthesis <- synthesis %>%
  select(columns_to_keep)

# Standardize column names
synthesis <- janitor::clean_names(synthesis)

# Print columns that are in both tables
intersect(names(synthesis), names(SAE_main))
```

Idem.

# Export final cleaned dataset ----

```{r}
write_csv2(SAE_main, "../data/intermediate/SAE_merged_2024.csv")
```

# Create a less exhaustive dataset ----

```{r}
# Keep only relevant columns for descriptive statistics
SAE_short_2024 <- SAE_main %>%
  select(
    # identification variables
    an, fi, rs, catr, stjr, rs, reg, dep, nomcom, cpo,
    # activity variables
    urg, regulation, passu, dt_hosp, stap, lit_uhcd,
    # urgences' personnel variables – number of hours
    hmed_tot, hmed_urg, hide_tot, hide_urg, hamb_tot, hmr_tot, hmramu_tot, arm_tot, osnp_tot, com_smur_su,
    # urgences' personnel variables – ETPs and effectifs
    etpsal_tot, etpsal_urg, efflib_tot, efflib_urg,
    # interns effectifs variables
    intmed, docju_med, inturg, docju_urg,
    # synthesis variables
    efflib_tot, etpsal_tot
  )
# Export less exhaustive dataset
write_csv2(SAE_short_2024, "../data/intermediate/SAE_merged_short_2024.csv")
```

