---
title: "SAE_analysis"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

# Set up ----

```{r setup, message=FALSE}
rm(list=ls())

path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
library(ggplot2)
library(sf)
library(viridis)
library(scales)
library(patchwork)
library(rlang)
library(gt)
```

# Import datasets ----

```{r cars, message=FALSE}
SAE_2024 <- read_csv2("../data/processed/SAE_final_2024.csv")
Legend <- read_csv2("../docs/SAE_2024/Legends_short.csv")
```


# 1. General discription of the dataset ----

To have a better overview of the structure of the dataset, we conduct descriptive statistics regarding establishment typology: frequencies/percentages for juridical_status, capacity, regulation, and emergency types.

```{r}
# Compute the percentage of emergency services by juridical status
status_counts <- SAE_2024 %>%
  count(juridical_status) %>%
  mutate(percentage = n / sum(n) * 100)
status_counts

# Compute the percentage of emergency services by capacity
capacity_counts <- SAE_2024 %>%
  count(capacity) %>%
  mutate(percentage = n / sum(n) * 100)
capacity_counts

# Compute the percentage of emergency services by regulation
regulation_counts <- SAE_2024 %>%
  count(regulation) %>%
  mutate(
    percentage = n / sum(n) * 100,
    regulation = case_when(
      regulation == 3 ~ "Yes, all the time",
      regulation == 2 ~ "Yes, part of the time",
      regulation == 1 ~ "No"
    )
  ) %>%
  rename("Is the emergency service regulated?" = regulation)
regulation_counts

# Compute the percentage of emergency services by type
type_counts <- SAE_2024 %>%
  count(emergency_type) %>%
  mutate(
    percentage = n / sum(n) * 100,
    emergency_type = case_when(
      emergency_type == 1 ~ "Pediatric",
      emergency_type == 2 ~ "General",
      emergency_type == 3 ~ "Both"
    )
  ) %>%
  rename("Type of emergency service" = emergency_type)
type_counts
```

We map those proportions into pie charts.

1. Status

```{r, echo = FALSE}
threshold <- 10  # percentage threshold for inside labels

status_plot <- status_counts %>%
  mutate(prop = percentage / 100) %>%
  arrange(desc(juridical_status)) %>%
  mutate(
    ymax = cumsum(prop),
    ymin = lag(ymax, default = 0),
    labelPosition = (ymax + ymin) / 2,
    label_inside = percentage >= threshold,
    
    # Define which slices are dark (adjust if needed)
    dark_slice = juridical_status %in% "Public"
  )

status_chart <- ggplot(status_plot) +
  
  geom_rect(aes(ymax = ymax, ymin = ymin,
                xmax = 1, xmin = 0,
                fill = juridical_status),
            color = "white") +
  
  coord_polar(theta = "y") +
  
  # Custom colors (NA in grey)
  scale_fill_manual(values = c(
    "Public" = "#440154",                 # dark purple
    "Privé non lucratif" = "#21908C",  # teal
    "Privé lucratif" = "#FDE725" # yellow
  )) +
  
  xlim(c(0, 1.35)) +
  
  # Leader lines for small slices
  geom_segment(
    data = dplyr::filter(status_plot, !label_inside),
    aes(x = 1, xend = 1.12,
        y = labelPosition, yend = labelPosition),
    color = "black",
    linewidth = 0.6
  ) +
  
    # Outside labels
  geom_text(
    data = dplyr::filter(status_plot, !label_inside),
    aes(x = 1.15, y = labelPosition,
        label = paste0(round(percentage, 0), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +

  # Inside labels
  geom_text(
    data = dplyr::filter(status_plot, label_inside),
    aes(x = 0.6, y = labelPosition,
        label = paste0(round(percentage, 0), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +
  
  scale_color_manual(values = c(
    "TRUE" = "white",   # white text for dark slices
    "FALSE" = "black"   # black text otherwise
  )) +
  
  labs(title = "Juridical Status of the Emergency Services",
       fill = "Status") +
  
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

status_chart
```

2. Type

```{r, echo = FALSE}
library(dplyr)
library(ggplot2)
library(viridis)
library(scales)

type_counts_plot <- type_counts %>%
  mutate(prop = percentage / 100) %>%
  arrange(desc(`Type of emergency service`)) %>%
  mutate(
    ymax = cumsum(prop),
    ymin = lag(ymax, default = 0),
    labelPosition = (ymax + ymin) / 2,
    label_inside = percentage >= threshold
  )

# ---- Compute viridis fill colors in the same order as the factor levels ----
type_levels <- type_counts_plot$`Type of emergency service`

fill_cols <- viridis::viridis(
  n = length(type_levels),
  option = "D",
  direction = 1
)
names(fill_cols) <- type_levels

# ---- Compute luminance to decide if text should be white or black ----
# Convert hex -> RGB in [0,1], compute relative luminance
hex_to_luminance <- function(hex) {
  rgb <- grDevices::col2rgb(hex) / 255
  r <- rgb[1, ]
  g <- rgb[2, ]
  b <- rgb[3, ]
  0.2126 * r + 0.7152 * g + 0.0722 * b
}

lum <- hex_to_luminance(fill_cols)

# Threshold: lower luminance = darker color => use white text
dark_levels <- names(lum[lum < 0.55])

type_counts_plot <- type_counts_plot %>%
  mutate(
    text_col = ifelse(`Type of emergency service` %in% dark_levels, "white", "black")
  )

type_chart <- ggplot(type_counts_plot) +
  # pie
  geom_rect(aes(
    ymax = ymax, ymin = ymin, xmax = 1, xmin = 0,
    fill = `Type of emergency service`
  ), color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = fill_cols) +
  xlim(c(0, 1.35)) +  # room for outside labels

  # leader lines ONLY for small slices
  geom_segment(
    data = dplyr::filter(type_counts_plot, !label_inside),
    aes(x = 1, xend = 1.12, y = labelPosition, yend = labelPosition),
    color = "black",
    linewidth = 0.6
  ) +

  # outside labels ONLY for small slices (keep black, more readable)
  geom_text(
    data = dplyr::filter(type_counts_plot, !label_inside),
    aes(x = 1.2, y = labelPosition,
        label = paste0(round(percentage, 1), "%")),
    size = 4,
    color = "black"
  ) +

  # inside labels ONLY for large slices (white if slice is dark)
  geom_text(
    data = dplyr::filter(type_counts_plot, label_inside),
    aes(x = 0.6, y = labelPosition,
        label = paste0(round(percentage, 1), "%"),
        color = text_col),
    size = 4,
    show.legend = FALSE
  ) +
  scale_color_identity() +

  labs(title = "Type of the Emergency Service", fill = "Type") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

type_chart

```

3. Regulation

```{r, include = FALSE}
regulation_plot <- regulation_counts %>%
  mutate(prop = percentage / 100) %>%
  arrange(desc(`Is the emergency service regulated?`)) %>%
  mutate(
    ymax = cumsum(prop),
    ymin = lag(ymax, default = 0),
    labelPosition = (ymax + ymin) / 2,
    label_inside = percentage >= threshold,
    
    # Define which slices are dark (adjust if needed)
    dark_slice = `Is the emergency service regulated?` %in% c("No", "Yes, all the time")
  )

regulation_chart <- ggplot(regulation_plot) +
  
  geom_rect(aes(ymax = ymax, ymin = ymin,
                xmax = 1, xmin = 0,
                fill = `Is the emergency service regulated?`),
            color = "white") +
  
  coord_polar(theta = "y") +
  
  # Custom colors (NA in grey)
  scale_fill_manual(values = c(
    "No" = "#440154",                 # dark purple
    "Yes, all the time" = "#21908C",  # teal
    "Yes, part of the time" = "#FDE725", # yellow
    "NA" = "grey70"
  )) +
  
  xlim(c(0, 1.35)) +
  
  # Leader lines for small slices
  geom_segment(
    data = dplyr::filter(regulation_plot, !label_inside),
    aes(x = 1, xend = 1.12,
        y = labelPosition, yend = labelPosition),
    color = "black",
    linewidth = 0.6
  ) +
  
  # Outside labels
  geom_text(
    data = dplyr::filter(regulation_plot, !label_inside),
    aes(x = 1.15, y = labelPosition,
        label = paste0(round(percentage, 1), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +
  
  # Inside labels
  geom_text(
    data = dplyr::filter(regulation_plot, label_inside),
    aes(x = 0.6, y = labelPosition,
        label = paste0(round(percentage, 1), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +
  
  scale_color_manual(values = c(
    "TRUE" = "white",   # white text for dark slices
    "FALSE" = "black"   # black text otherwise
  )) +
  
  labs(title = "Regulation of the Emergency Services",
       fill = "Regulation") +
  
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

regulation_chart
```

4. Capacity

```{r}
capacity_plot_2 <- capacity_counts %>%
  mutate(prop = percentage / 100) %>%
  arrange(desc(capacity)) %>%
  mutate(
    ymax = cumsum(prop),
    ymin = lag(ymax, default = 0),
    labelPosition = (ymax + ymin) / 2,
    label_inside = percentage >= threshold,
    
    # Define which slices are dark (adjust if needed)
    dark_slice = capacity %in% "High capacity"
  )

capacity_chart <- ggplot(capacity_plot_2) +
  
  geom_rect(aes(ymax = ymax, ymin = ymin,
                xmax = 1, xmin = 0,
                fill = capacity),
            color = "white") +
  
  coord_polar(theta = "y") +
  
  # Custom colors (NA in grey)
  scale_fill_manual(values = c(
    "High capacity" = "#440154",                 # dark purple
    "Medium capacity" = "#21908C",  # teal
    "Low capacity" = "#FDE725" # yellow
  )) +
  
  xlim(c(0, 1.35)) +
  
  # Leader lines for small slices
  geom_segment(
    data = dplyr::filter(capacity_plot_2, !label_inside),
    aes(x = 1, xend = 1.12,
        y = labelPosition, yend = labelPosition),
    color = "black",
    linewidth = 0.6
  ) +
  
  # Outside labels
  geom_text(
    data = dplyr::filter(capacity_plot_2, !label_inside),
    aes(x = 1.15, y = labelPosition,
        label = paste0(round(percentage, 1), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +
  
  # Inside labels
  geom_text(
    data = dplyr::filter(capacity_plot_2, label_inside),
    aes(x = 0.6, y = labelPosition,
        label = paste0(round(percentage, 0), "%"),
        color = dark_slice),
    size = 4,
    show.legend = FALSE
  ) +
  
  scale_color_manual(values = c(
    "TRUE" = "white",   # white text for dark slices
    "FALSE" = "black"   # black text otherwise
  )) +
  
  labs(title = "Capacity of the Emergency Services",
       fill = "Capacity") +
  
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

capacity_chart
```

```{r}
# Export the 4 pie charts
ggsave("../results/SAE_2024/status_pie_chart.png", plot = status_chart, width = 6, height = 6)
ggsave("../results/SAE_2024/type_pie_chart.png", plot = type_chart, width = 6, height = 6)
ggsave("../results/SAE_2024/regulation_pie_chart.png", plot = regulation_chart, width = 6, height = 6)
ggsave("../results/SAE_2024/capacity_pie_chart.png", plot = capacity_chart, width = 6, height = 6)
```

```{r}
# Display the 4 charts together
(status_chart + type_chart) / (regulation_chart + capacity_chart) +
  plot_annotation(
                  theme = theme(plot.title = element_text(hjust = 0.5, size = 16)))
```

```{r}
rm(status_counts, type_counts, regulation_counts, capacity_counts)
rm(status_plot, type_counts_plot, regulation_plot, capacity_plot_2)
rm(status_chart, type_chart, regulation_chart, capacity_chart)
```

We then check for correlation between capacity and visits (scatter plot): hospitals with high capacity and high number of visits are just giants, the ones with lots of visits but low capacity are overloaded (it introduces the 2nd part).

```{r}
# Calculate medians first (optional, but cleaner)
median_beds <- median(SAE_2024$beds_number, na.rm = TRUE)
median_visits <- median(SAE_2024$total_emergency_visits_win, na.rm = TRUE)

ggplot(SAE_2024, aes(x = beds_number, y = total_emergency_visits_win)) +
  geom_point(alpha = 0.6) +
  # Add the quadrant lines
  geom_vline(xintercept = median_beds, linetype = "dashed", color = "red", size = 0.8) +
  geom_hline(yintercept = median_visits, linetype = "dashed", color = "red", size = 0.8) +
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Correlation between Capacity and Number of Visits",
    subtitle = "Quadrants divided by median values",
    x = "Capacity (number of beds)",
    y = "Number of Visits in Emergency Services"
  ) +
  theme_minimal()

ggsave("../results/SAE_2024/capacity_visits_scatter.png", width = 10, height = 8)
```

We focus on those in the top left quadrant: low capacity but high number of visits. We check if they are public or private, and if they are regulated or not.

```{r}
# Filter for the top-left quadrant
overloaded_hosp <- SAE_2024 %>%
  filter(beds_number < median_beds, 
         total_emergency_visits_win > median_visits)

# Define is_overloaded variable (top-left corner of the scatterplot) in SAE_2024
threshold_beds <- median(SAE_2024$beds_number, na.rm = TRUE)
threshold_visits <- median(SAE_2024$total_emergency_visits_win, na.rm = TRUE)
SAE_2024 <- SAE_2024 %>%
  mutate(is_overloaded = if_else(
    beds_number < threshold_beds & total_emergency_visits_win > threshold_visits, 
    "Overloaded", 
    "Not Overloaded"
  ))
```

```{r}
# See what percentage of them are public
overloaded_hosp %>%
  count(juridical_status) %>%
  mutate(percentage_status = n / sum(n) * 100)

# See what percentage of them are regulated
overloaded_hosp %>%
  count(regulation) %>%
  mutate(
    percentage = n / sum(n) * 100,
    regulation = case_when(
      regulation == 3 ~ "Yes, all the time",
      regulation == 2 ~ "Yes, part of the time",
      regulation == 1 ~ "No"
    )
  ) %>%
  rename("Is the emergency service regulated?" = regulation)

# See what percentage of them are pediatric/general/both
overloaded_hosp %>%
  count(emergency_type) %>%
  mutate(
    percentage = n / sum(n) * 100,
    emergency_type = case_when(
      emergency_type == 1 ~ "Pediatric",
      emergency_type == 2 ~ "General",
      emergency_type == 3 ~ "Both"
    )
  ) %>%
  rename("Type of emergency service" = emergency_type)
```

The overloaded hospitals are mostly public (48%), mostly regulated (48%), and mostly general (98%). But Chi-squared tests.

We check that the distributions are statistically significant.

```{r}
# Create function to compute the distribution of overloaded vs non-overloaded hospitals for a given variable, and run a chi-squared test
analyze_overload_dist <- function(data, var_name) {
  var_sym <- sym(var_name)
  
  global_dist <- data %>%
    count(!!var_sym) %>%
    mutate(pct_global = n / sum(n) * 100) %>%
    select(!!var_sym, pct_global)

  overloaded_dist <- data %>%
    filter(is_overloaded == "Overloaded") %>%
    count(!!var_sym) %>%
    mutate(pct_overloaded = n / sum(n) * 100) %>%
    # right_join ensures we keep categories that have 0 occurrences
    right_join(global_dist, by = var_name) %>%
    mutate(
      pct_overloaded = replace_na(pct_overloaded, 0)
    )

  table_stats <- data %>%
    count(is_overloaded, !!var_sym) %>%
    pivot_wider(names_from = all_of(var_name), values_from = n, values_fill = 0) %>%
    select(-is_overloaded) %>%
    as.matrix()
  
  chi_result <- chisq.test(table_stats)

  return(list(distribution_table = overloaded_dist, chi_test = chi_result))
}

# Define variables to test
vars_to_test <- c("emergency_type", "regulation", "juridical_status")

```

```{r}
# Create contingency tables and run chi-squared tests for each variable
results <- setNames(lapply(vars_to_test, function(v) analyze_overload_dist(SAE_2024, v)), vars_to_test)
results
```
```{r}
# Create a summary table for the distribution of overloaded vs non-overloaded hospitals for each variable
chi2_summary_table <- imap_dfr(results, function(res, var_name) {
  res$distribution_table %>%
    mutate(
      variable = var_name,
      category_code = as.character(!!sym(var_name)), # Keep code for mapping
      p_value = res$chi_test$p.value
    ) %>%
    select(variable, category_code, pct_overloaded, pct_global, p_value)
})

# Add the labels
chi2_summary_table <- chi2_summary_table %>%
  mutate(
    category_label = case_when(
      variable == "emergency_type" & category_code == "1" ~ "Pediatric",
      variable == "emergency_type" & category_code == "2" ~ "General",
      variable == "emergency_type" & category_code == "3" ~ "Both",
      variable == "regulation" & category_code == "1" ~ "No",
      variable == "regulation" & category_code == "2" ~ "Part-time",
      variable == "regulation" & category_code == "3" ~ "Full-time",
      variable == "juridical_status" & category_code == "Public" ~ "Public",
      variable == "juridical_status" & category_code == "Privé non lucratif" ~ "Private non-profit",
      variable == "juridical_status" & category_code == "Privé lucratif" ~ "Private for-profit",
      TRUE ~ category_code 
    )
  )

# Calculate the difference in percentage points between the overloaded distribution and the global distribution
chi2_summary_table <- chi2_summary_table %>%
  mutate(pct_diff = pct_overloaded - pct_global)

chi2_summary_table <-  chi2_summary_table %>%
  select(variable, category_label, pct_overloaded, pct_global, pct_diff, p_value)

chi2_summary_table
```


```{r}
threshold <- quantile(SAE_2024$patients_per_doctor_win, 0.9, na.rm = TRUE)

# Run the analysis for each variable and store results in a list
all_results <- setNames(lapply(vars_to_test, function(v) analyze_overload_dist(SAE_2024, v)), vars_to_test)

# Compile the single summary table with labels
summary_table <- imap_dfr(all_results, function(res, var_name) {
  res$distribution_table %>%
    mutate(
      variable = var_name,
      category_code = as.character(!!sym(var_name)), # Keep code for mapping
      p_value = res$chi_test$p.value
    ) %>%
    select(variable, category_code, pct_overloaded, pct_global, p_value)
})

# Map the labels and add the difference column
summary_table_final <- summary_table %>%
  mutate(
    # Map labels to codes
    category_label = case_when(
      variable == "emergency_type" & category_code == "1" ~ "Pediatric",
      variable == "emergency_type" & category_code == "2" ~ "General",
      variable == "emergency_type" & category_code == "3" ~ "Both",
      variable == "regulation" & category_code == "1" ~ "No",
      variable == "regulation" & category_code == "2" ~ "Part-time",
      variable == "regulation" & category_code == "3" ~ "Full-time",
      TRUE ~ category_code 
    ),
    # Compute the Difference (Percentage Points)
    pct_diff = pct_overloaded - pct_global,
    
    # Clean up formatting
    across(starts_with("pct"), ~round(., 2)),
    p_value = format.pval(p_value, digits = 3, eps = 0.001),
    
    # Add interpretation column
    representation = ifelse(pct_diff > 0, "Over-represented", "Under-represented")
  ) %>%
  select(variable, category_label, pct_overloaded, pct_global, pct_diff, p_value, representation)

# Display the final summary table
summary_table_final
```

```{r}
# Export the summary table as a PNG image
summary_table_final %>%
  gt() %>%
  tab_header(title = "Distribution of the Overloaded Hospitals") %>%
  gtsave("../results/SAE_2024/overload_distribution_summary.png")
```

# 2. Ratios of doctors per inhabitant, and visits per doctor ----

We first want to understand the distribution of doctors across France, to identify potentially underserved areas. We compute the ratio of patients per doctor by department, and map it.

The population data is from Insee.

## 2.1. Doctors per inhabitant

```{r, include = FALSE}
# Import demographic data
demographics <- read_csv2("../data/raw/pop_dpt.csv")

# Keep only columns 1, 2 and 8 and rows 5 to 108
demographics <- demographics %>%
  select(1, 2, 8) %>%
  slice(5:108) %>%
  rename(department_code = `Estimation de population au 1er janvier, par département, sexe et grande classe d'âge`) %>%
  rename(department_name = 2) %>%
  rename(population = 3)

# Rename oversees territories department code to match the one in the main dataset
demographics <- demographics %>%
  mutate(department = case_when(
    department_code == "971" ~ "9A", # Guadeloupe
    department_code == "972" ~ "9B", # Martinique
    department_code == "973" ~ "9C", # Guyane
    department_code == "974" ~ "9D", # La Réunion
    department_code == "976" ~ "9F", # Mayotte
    TRUE ~ department_code # Keeps everything else as is
  )) %>%
  select(-department_code) %>%
  rename(department_code = department)

# Force to numeric
demographics <- demographics %>%
  mutate(
    population = as.numeric(str_replace(population, ",", "."))
  )

# Drop rows 97, 103 and 104
demographics <- demographics %>%
  slice(-c(97, 103, 104))
```

```{r}
# Add the number of liberal physicians
SAE_2024 <- SAE_2024 %>%
  mutate(
    # Convert liberals number to ETP
    etp_liberals = number_physicians_liberal,
    # Sum the ETP of employed physicians and liberals
    etp_total_combine = etp_physicians_empl_emergency + etp_liberals,
  )
```

```{r}
# Convert department codes to characters
SAE_2024 <-  SAE_2024 %>%
  mutate(dep_number = as.character(dep_number))

# Sum the etp_emergency_win by department
doctors_by_department <- SAE_2024 %>%
  group_by(dep_number) %>%
  summarise(total_etp_emergency = sum(etp_total_combine, na.rm = TRUE))

# # Idem for liberal doctors
# liberals_by_department <- SAE_2024 %>%
#   group_by(dep_number) %>%
#   summarise(total_liberal_doctors = sum(number_physicians_liberal, na.rm = TRUE))

# Merge the number of doctors with the demographic data
doctors_pop <- doctors_by_department %>%
  left_join(demographics, by = c("dep_number" = "department_code"))

# #Merge with the liberal doctors data
# doctors_pop <- doctors_pop %>%
#   left_join(liberals_by_department, by = "dep_number")

# Calculate the ratio of doctors per 1000 inhabitant (order by ascending ratio)
doctors_pop <- doctors_pop %>%
  mutate(doctors_per_1000 = (total_etp_emergency / population) * 1000)
  # mutate(liberals_per_1000 = (total_liberal_doctors / population) * 1000)

#Check for correlation
# cor(doctors_pop$doctors_per_1000, doctors_pop$liberals_per_1000, use = "complete.obs")
```

Negative correlation => the more emergency doctors there are in a department, the fewer liberal doctors.

We create a map of the ratio of doctors per 1000 inhabitants by department, to identify potential disparities in access to emergency care across France.

```{r}
# Load shapefile for French departments
departments_geo <- st_read("../data/raw/departements.geojson")

# Merge the geographic data with the doctors per population data
departments_geo <- departments_geo %>%
  left_join(doctors_pop, by = c("code" = "dep_number"))

# Choropleth map of doctors per 1000 inhabitants by department (scale_fill_viridis_c)
doctors_per_1000 <- ggplot(data = departments_geo) +
  geom_sf(aes(fill = doctors_per_1000), color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Doctors per 1000 inhabitants",
    na.value = "grey90"
  ) +
  labs(title = "Ratio of Doctors in Emergency Services per 1000 Inhabitants by Department") +
  theme_minimal()

# Choropleth map of liberal doctors per 1000 inhabitants by department (scale_fill_viridis_d)
liberals_per_1000 <- ggplot(data = departments_geo) +
  geom_sf(aes(fill = liberals_per_1000), color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Liberal Doctors per 1000 inhabitants",
    na.value = "grey90"
  ) +
  labs(title = "Ratio of Liberal Doctors in Emergency Services per 1000 Inhabitants by Department") +
  theme_minimal()
```

```{r}
# Export the maps
ggsave("../results/SAE_2024/doctors_per_1000_map.png", doctors_per_1000, width = 10, height = 8)
#ggsave("../results/SAE_2024/liberals_per_1000_map.png", liberals_per_1000, width = 10, height = 8)
```

--

TO REFLECT ON: not sure it's useful to explore le fait que la où il y a peu de médecins par habitants les établissements sont publics.

We compare this ratio across juridical status and regulation dimensions.

```{r}
# Merge the index of doctors per 1000 inhabitants with the main dataset
SAE_2024 <- SAE_2024 %>%
  left_join(doctors_pop %>% select(dep_number, doctors_per_1000), by = c("dep_number" = "dep_number")) %>%
  rename(dep_doctors_per_1000 = doctors_per_1000)

# Boxplot of doctors per 1000 inhabitants by juridical status
ggplot(SAE_2024, aes(x = juridical_status, y = dep_doctors_per_1000)) +
  geom_boxplot() +
  labs(title = "Distribution of Doctors per 1000 Inhabitants by Juridical Status",
       x = "Juridical Status",
       y = "Doctors per 1000 Inhabitants") +
  theme_minimal()
```

```{r}
# Compute the average index of doctors per 1000 inhabitants by department, and the percentage of private hospitals in each department
dep_analysis <- SAE_2024 %>%
  group_by(dep_number) %>%
  summarise(
    total_hospitals = n(),
    pct_private = sum(juridical_status %in% c("Privé lucratif", "Privé non lucratif")) / n() * 100,
    index_urgences = mean(dep_doctors_per_1000, na.rm = TRUE)
  )

# Scatter plot
ggplot(dep_analysis, aes(x = pct_private, y = index_urgences)) +
  geom_point(aes(size = total_hospitals), alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", color = "red", se = FALSE, linetype = "dashed") +
  labs(
    title = "Correlation between the Percentage of Private Hospitals and the Index of Doctors per 1000 Inhabitants",
    subtitle = "Each point represents a department, size of the points corresponds to the number of hospitals in the department",
    x = "Percentage of Private Hospitals in the Department",
    y = "Index of Doctors per 1000 Inhabitants",
    size = "Number of Hospitals in the Department"
  ) +
  theme_minimal()
```

## 2.2. Visits per doctor

As mentioned in the data cleaning code on outliers, we use the raw variables in these calculations (bc the outliers didn't appear to be the result of a filling mistake).

```{r}
# Calculate the ratio of visits per doctor EMPLOYED
SAE_2024 <- SAE_2024 %>% 
  mutate(patients_per_doctor = case_when(
    etp_physicians_empl_emergency == 0 ~ NA_real_,
    is.na(etp_physicians_empl_emergency) ~ NA_real_,
    TRUE ~ total_emergency_visits_raw / etp_physicians_empl_emergency
  ))

# Compute the ratio of visits per doctor employed AND LIBERAL
SAE_2024 <- SAE_2024 %>%
  mutate(
    patients_per_doctor_combined = total_emergency_visits_raw / etp_total_combine
  )

# Replace outliers by NA
SAE_2024 <- SAE_2024 %>%
  mutate(patients_per_doctor_clean = case_when(
      patients_per_doctor_combined > 10000 ~ NA_real_,  # Too high
      patients_per_doctor_combined < 300   ~ NA_real_,  # Too low
      is.infinite(patients_per_doctor_combined) ~ NA_real_,
      TRUE ~ patients_per_doctor_combined
    )
  )

# Rename the original patients_per_doctor variable to keep it for comparison, and rename the cleaned one to patients_per_doctor
SAE_2024 <-  SAE_2024 %>%
  rename(patients_per_doctor_empl = patients_per_doctor)
SAE_2024 <-  SAE_2024 %>%
  rename(patients_per_doctor = patients_per_doctor_clean)
```

```{r}
# Create a dataset with the average ratio of patients per doctor by department
ratio_by_department <- SAE_2024 %>%
  group_by(dep_number) %>%
  summarise(avg_patients_per_doctor = mean(patients_per_doctor, na.rm = TRUE))
```

HOPITAL PRIVE DE VITRY SITE PASTEUR is an obvious outlier (624950.000 patients per doctors, which would mean 1,700 per doctor per day). It's likely due to the low value of its etp_physicians_empl_emergency (0.01), which is probably a filling mistake. We winsorize the patients_per_doctor variable at 99% to limit the influence of this outlier on the average ratio by department.


While the primary variables were winsorized at the 95th percentile to account for the high capacity of University Hospitals (CHUs), the resulting visits-per-doctor ratio still exhibited extreme outliers driven by 'denominator effects.'

In several cases, such as Hôpital Privé de Vitry, APHM Hôpital La Timone, and Institut Arnault Tzanck, the recorded ETP is exceptionally low (e.g., 0.04 to 0.11), resulting in mathematically impossible ratios exceeding 100,000 visits per physician per year. Similar artifacts are observed for GCS Urgences Pays Royannais, Polyclinique du Beaujolais, and Clinique les Ormeaux. These values likely stem from reporting discrepancies in the SAE database—such as the use of external contractors, interim staff, or medical personnel being attached to different legal entities—rather than reflecting actual clinical throughput.

To preserve these observations while preventing these 'artificial' efficiencies from biasing the regression coefficients, we winsorize the ratio at the 99th percentile. This approach retains the high-activity nature of these centers but caps their impact at a statistically plausible upper bound.

```{r}
# Winsorize the patients_per_doctor variable at 99%
SAE_2024 <- SAE_2024 %>%
  mutate(patients_per_doctor_win = ifelse(
    patients_per_doctor > quantile(patients_per_doctor, 0.99, na.rm = TRUE),
    quantile(patients_per_doctor, 0.99, na.rm = TRUE),
    patients_per_doctor
  ))

# Update the ratio_by_department dataset with the winsorized variable
rm(ratio_by_department)
ratio_by_department <- SAE_2024 %>%
  group_by(dep_number) %>%
  summarise(avg_patients_per_doctor = mean(patients_per_doctor_win, na.rm = TRUE))
```

```{r}
# Merge the updated dataset with the geographic data
ratio_by_department <- departments_geo %>%
  left_join(ratio_by_department, by = c("code" = "dep_number"))

# Choropleth map of the average ratio of patients per doctor by department
patients_per_doctor <- ggplot(data = ratio_by_department) +
  geom_sf(aes(fill = avg_patients_per_doctor), color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Average Patients per Doctor",
    na.value = "grey90"
  ) +
  labs(title = "Average Ratio of Patients per Doctor in Emergency Services by Department") +
  theme_minimal()

patients_per_doctor
```

```{r}
ggsave("../results/SAE_2024/patients_per_doctor_map.png",  patients_per_doctor, width = 10, height = 8)
```

We now compare the ratio across juridical status and regulation dimensions.

```{r}
# Compute the median ratio of patients per doctor
national_median <- median(SAE_2024$patients_per_doctor_win, na.rm = TRUE)

#Prepare the data for Juridical Status
status_deviation <- SAE_2024 %>%
  group_by(juridical_status) %>%
  summarise(group_median = median(patients_per_doctor_win, na.rm = TRUE)) %>%
  mutate(
    deviation = group_median - national_median,
    type = ifelse(deviation > 0, "Above Median", "Below Median")
  ) %>%
  arrange(deviation)

# 3. Plot for Juridical Status
status_dev_plot <- ggplot(status_deviation, aes(x = reorder(juridical_status, deviation), y = deviation, fill = type)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("Above Median" = "#d73027", "Below Median" = "#4575b4")) +
  labs(
    title = "Deviation from National Median by Juridical Status",
    subtitle = paste("National Median:", round(national_median, 1), "visits per doctor"),
    x = "Juridical Status",
    y = "Difference from National Median",
    fill = "Status"
  ) +
  theme_minimal()

status_dev_plot
```

```{r}
# Prepare data for capacity
capacity_deviation <- SAE_2024 %>%
  group_by(capacity) %>%
  summarise(group_median = median(patients_per_doctor_win, na.rm = TRUE)) %>%
  mutate(
    deviation = group_median - national_median,
    type = ifelse(deviation > 0, "Above Median", "Below Median"),
    capacity_label = case_when(
      capacity == 3 ~ "High capacity",
      capacity == 2 ~ "Medium capacity",
      capacity == 1 ~ "Low capacity"
    )
  ) %>%
  arrange(deviation)

# Plot for capacity
capacity_dev_plot <- ggplot(capacity_deviation, aes(x = reorder(capacity, deviation), y =
    deviation, fill = type)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("Above Median" = "#d73027", "Below Median" = "#4575b4")) +
  labs(
    title = "Deviation from National Median by Capacity",
    subtitle = paste("National Median:", round(national_median, 1), "visits per doctor"),
    x = "Capacity",
    y = "Difference from National Median",
    fill = "Capacity"
  ) +
  theme_minimal()

capacity_dev_plot
```


```{r}
# Prepare data for regulation
regulation_deviation <- SAE_2024 %>%
  group_by(regulation) %>%
  summarise(group_median = median(patients_per_doctor_win, na.rm = TRUE)) %>%
  mutate(
    deviation = group_median - national_median,
    type = ifelse(deviation > 0, "Above Median", "Below Median"),
    regulation_label = case_when(
      regulation == 3 ~ "Yes, all the time",
      regulation == 2 ~ "Yes, part of the time",
      regulation == 1 ~ "No"
    )
  ) %>%
  arrange(deviation)

# Plot for regulation
regulation_dev_plot <- ggplot(regulation_deviation, aes(x = reorder(regulation_label, deviation), y =
    deviation, fill = type)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("Above Median" = "#d73027", "Below Median" = "#4575b4")) +
  labs(
    title = "Deviation from National Median by Regulation Status",
    subtitle = paste("National Median:", round(national_median, 1), "visits per doctor"),
    x = "Regulation Status",
    y = "Difference from National Median",
    fill = "Regulation"
  ) +
  theme_minimal()

regulation_dev_plot
```

```{r}
# Prepare data for waiting_monitoring
waiting_deviation <- SAE_2024 %>%
  group_by(waiting_monitoring) %>%
  summarise(group_median = median(patients_per_doctor_win, na.rm = TRUE)) %>%
  mutate(
    deviation = group_median - national_median,
    type = ifelse(deviation > 0, "Above Median", "Below Median"),
  ) %>%
  arrange(deviation)

# Plot for waiting monitoring
waiting_dev_plot <- ggplot(waiting_deviation, aes(x = reorder(waiting_monitoring, deviation), y =
    deviation, fill = type)) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_manual(values = c("Above Median" = "#d73027", "Below Median" = "#4575b4")) +
  labs(
    title = "Deviation from National Average by Waiting Monitoring Status",
    subtitle = paste("National Average:", round(national_median, 1), "visits per doctor"),
    x = "Waiting Monitoring Status",
    y = "Difference from National Average",
    fill = "Waiting Monitoring"
  ) +
  theme_minimal()

waiting_dev_plot
```

```{r}
ggsave("../results/SAE_2024/status_deviation_plot.png", status_dev_plot, width = 8, height = 6)
ggsave("../results/SAE_2024/capacity_deviation_plot.png", capacity_dev_plot, width = 8, height = 6)
ggsave("../results/SAE_2024/regulation_deviation_plot.png", regulation_dev_plot, width = 8, height = 6)
ggsave("../results/SAE_2024/waiting_deviation_plot.png", waiting_dev_plot, width = 8, height = 6)
```

So now we've seen:
- the hospitals that are most overloaded in terms of visits per doctor, their deviation from the national median across the different dimensions (juridical status, regulation, capacity, waiting monitoring), and where there are mostly located,
- the departments with the lowest ratio of doctors per inhabitant.

What we are trying to explain is why some hospitals are more overloaded than others. We have seen that the regulation and capacity dimensions don't seem to be correlated with the ratio of patients per doctor, but the juridical status does (private hospitals have a lower ratio of patients per doctor). We also know that some departments have a very low ratio of doctors per inhabitant.


```{r}
# Scatterplot of patients per doctor vs. doctors per 1000 inhabitants by department (with a regression line)
ggplot(ratio_by_department, aes(x = doctors_per_1000, y = avg_patients_per_doctor)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Correlation between Doctors per 1000 Inhabitants and Average Patients per Doctor",
    x = "Doctors per 1000 Inhabitants",
    y = "Average Patients per Doctor"
  ) +
  theme_minimal()

cor(ratio_by_department$doctors_per_1000, ratio_by_department$avg_patients_per_doctor, use = "complete.obs")
```
Negative correlation: the more doctors there are in a department, the fewer patients per doctor there are on average.

```{r}
# Save the scatter plot
ggsave("../results/SAE_2024/patients_per_doctor_vs_doctors_per_1000.png", width = 10, height = 8)
```


```{r}
# Statistical test on the correlation
cor_test <- cor.test(ratio_by_department$doctors_per_1000, ratio_by_department$avg_patients_per_doctor, use = "complete.obs")
print(cor_test)
```
We reject the null hypothesis at the 1% threshold: the correlation is statistically significant. However, it is moderate (-0.35), which means that there are other factors at play in the ratio of patients per doctor.

Moreover, the R^2 is equal to 0.12, which means that only 12% of the variance in the ratio of patients per doctor is explained by the number of doctors per inhabitant.

We draw the same scatterplot but this time at the hospital level. We focus on the hospitals in the top-left corner, ie those with a low ratio of doctors per inhabitant but a high ratio of patients per doctor, to see if they are mostly public or private, and if they are regulated or not.

```{r}
# Scatter plot of patients per doctor vs. doctors per 1000 inhabitants by hospital (divide in 4 quadrants based on the median)
ggplot(SAE_2024, aes(x = dep_doctors_per_1000, y = patients_per_doctor_win)) +
  geom_point(alpha = 0.6) +
  geom_vline(xintercept = median(SAE_2024$dep_doctors_per_1000, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  geom_hline(yintercept = median(SAE_2024$patients_per_doctor_win, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  labs(
    title = "Correlation between Doctors per 1000 Inhabitants and Patients per Doctor at the Hospital Level",
    subtitle = "Quadrants divided by median values",
    x = "Doctors per 1000 Inhabitants in the Department",
    y = "Patients per Doctor (winsorized)"
  ) +
  theme_minimal()
```
```{r}
ggsave("../results/SAE_2024/patients_per_doctor_vs_doctors_per_1000_hospital.png", width = 10, height = 8)
```

We study the top left corner, ie the departments with low doctors per inhabitant but high patients per doctor. We check if they are mostly public or private, and if they are regulated or not.

NB: We add the percentages on the whole dataset to have a benchmark for comparison (eg if 80% of the hospitals in the top left corner are public but only 50% of the whole dataset is public, then we can say that public hospitals are overrepresented in the top left corner). We test this difference with a Chi squared test.

```{r}
# Filter for the top-left quadrant
tensed_hospitals <- SAE_2024 %>%
  filter(dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE), 
         patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE))
```

```{r}
# # Create the "is_tensed" column
# SAE_2024 <- SAE_2024 %>%
#   mutate(is_tensed = ifelse(
#     dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE) & 
#     patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE),
#     "Tensed", "Not Tensed"
#   ))
# 
# # Run the analysis
# all_results <- setNames(lapply(vars_to_test, function(v) analyze_overload_dist(SAE_2024, v)), vars_to_test)
# 
# # Compile the single summary table with Labels
# summary_table <- imap_dfr(all_results, function(res, var_name) {
#   res$distribution_table %>%
#     mutate(
#       variable = var_name,
#       category_code = as.character(!!sym(var_name)), # Keep code for mapping
#       p_value = res$chi_test$p.value
#     ) %>%
#     select(variable, category_code, pct_overloaded, pct_global, p_value)
# })
# # Map the labels and add the Difference column
# summary_table_final <- summary_table %>%
#   mutate(
#     # 1. Map labels to codes
#     category_label = case_when(
#       variable == "emergency_type" & category_code == "1" ~ "Pediatric",
#       variable == "emergency_type" & category_code == "2" ~ "General",
#       variable == "emergency_type" & category_code == "3" ~ "Both",
#       variable == "regulation" & category_code == "1" ~ "No",
#       variable == "regulation" & category_code == "2" ~ "Part-time",
#       variable == "regulation" & category_code == "3" ~ "Full-time",
#       TRUE ~ category_code
#     ),
#     # 2. Compute the Difference (Percentage Points)
#     pct_diff = pct_overloaded - pct_global,
#     # 3. Clean up formatting
#     across(starts_with("pct"), ~round(., 2)),
#     p_value = format.pval(p_value, digits = 3, eps = 0.001),
#     # 4. Add a simple interpretation column
#     representation = ifelse(pct_diff > 0, "Over-represented", "Under-represented")
#   ) %>%
#   select(variable, category_label, pct_overloaded, pct_global, pct_diff, p_value, representation)
# 
# summary_table_final
```

```{r}
# Export the table
# summary_table_final %>%
#   gt() %>%
#   tab_header(title = "Distribution of the Tensed Hospitals") %>%
#   gtsave("../results/SAE_2024/tensed_hospitals_distribution.png")
```

1. Juridical status

```{r}
# Global reference
global_status <- SAE_2024 %>%
  count(juridical_status) %>%
  mutate(percentage_global = n / sum(n) * 100) %>%
  select(juridical_status, percentage_global)

# Table
tensed_hospitals %>%
  count(juridical_status) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_status, by = "juridical_status")
```

```{r}
# Create a binary variable for being in the tensed group
temp_status <- SAE_2024 %>%
  mutate(is_tensed = ifelse(
    dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE) & 
    patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE),
    "Tensed", "Not Tensed"
  ))

# Table of counts for the Chi-squared test
table_stats <- table(temp_status$is_tensed, SAE_2024$juridical_status)
table_stats

# Chi-squared test
prop.test(table_stats[, "Public"], rowSums(table_stats))
prop.test(table_stats[, "Privé lucratif"], rowSums(table_stats))
prop.test(table_stats[, "Privé non lucratif"], rowSums(table_stats))
```

No statistically significant difference between the proportion of public and private hospitals in the tensed group compared to the whole dataset.

2. Regulation

```{r}
# Global reference
global_reg <- SAE_2024 %>%
  count(regulation) %>%
  mutate(percentage_global = n / sum(n) * 100) %>%
  select(regulation, percentage_global)

# Table
tensed_hospitals %>%
  count(regulation) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_reg, by = "regulation") %>%
  mutate(regulation = case_when(
      regulation == 3 ~ "Yes, all the time",
      regulation == 2 ~ "Yes, part of the time",
      regulation == 1 ~ "No"
  ))
```


```{r}
# Create a binary variable for being in the tensed group
temp_reg <- SAE_2024 %>%
  mutate(is_tensed = ifelse(
    dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE) &
    patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE),
    "Tensed", "Not Tensed"
  ))

# Table of counts for the Chi-squared test
table_reg <- table(temp_reg$is_tensed, SAE_2024$regulation)
table_reg

# Chi-squared test
prop.test(table_reg[, "1"], rowSums(table_reg))
prop.test(table_reg[, "2"], rowSums(table_reg))
prop.test(table_reg[, "3"], rowSums(table_reg))

```

Same.

3. Type

```{r}
# Global reference
global_type <- SAE_2024 %>%
  count(emergency_type) %>%
  mutate(percentage_global = n / sum(n) * 100) %>%
  select(emergency_type, percentage_global)

# Table
tensed_hospitals %>%
  count(emergency_type) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_type, by = "emergency_type") %>%
  mutate(emergency_type = case_when(
      emergency_type == 1 ~ "Pediatric",
      emergency_type == 2 ~ "General",
      emergency_type == 3 ~ "Both"
  ))
```

```{r}
# Create a binary variable for being in the tensed group
temp_type <- SAE_2024 %>%
  mutate(is_tensed = ifelse(
    dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE) &
    patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE),
    "Tensed", "Not Tensed"
  ))

# Table of counts for the Chi-squared test
table_type <- table(temp_type$is_tensed, SAE_2024$emergency_type)
table_type

# Chi-squared test
prop.test(table_type[, "1"], rowSums(table_type))
prop.test(table_type[, "2"], rowSums(table_type))
prop.test(table_type[, "3"], rowSums(table_type))
```

Difference stat signif at the 5% threshold for general hospitals: they are overrepresented in the tensed group compared to the whole dataset.

4. Capacity

```{r}
# Global reference
global_cap <- SAE_2024 %>%
  count(capacity) %>%
  mutate(percentage_global = n / sum(n) * 100) %>%
  select(capacity, percentage_global)

# Table
tensed_hospitals %>%
  count(capacity) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_cap, by = "capacity")
```

```{r}
# Create a binary variable for being in the tensed group
temp_cap <- SAE_2024 %>%
  mutate(is_tensed = ifelse(
    dep_doctors_per_1000 < median(dep_doctors_per_1000, na.rm = TRUE) & 
    patients_per_doctor_win > median(patients_per_doctor_win, na.rm = TRUE),
    "Tensed", 
    "Not Tensed"
  ))

# Table of counts for the Chi-squared test
table_cap <- table(temp_cap$is_tensed, SAE_2024$capacity)
table_cap

# Chi-squared test
prop.test(table_cap[, "High capacity"], rowSums(table_cap))
prop.test(table_cap[, "Medium capacity"], rowSums(table_cap))
prop.test(table_cap[, "Low capacity"], rowSums(table_cap))
```

Difference stat signif at the 1% threshold for high capacity hospitals: they are overrepresented in the tensed group compared to the whole dataset.
Difference stat signif at the 1% threshold for low capacity hospitals: they are underrepresented in the tensed group compared to the whole dataset.

```{r}
# Create a table with the percentage_tensed and the percentage_global across the four variables
tensed_summary <- tensed_hospitals %>%
  count(juridical_status, regulation, emergency_type, capacity) %>%
  group_by(juridical_status) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_status, by = "juridical_status") %>%
  group_by(regulation) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_reg, by = "regulation") %>%
  group_by(emergency_type) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_type, by = "emergency_type") %>%
  group_by(capacity) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_cap, by = "capacity")

tensed_summary
```


--> Check if those are the same hospitals as the ones we identified as overloaded in the first part (high number of visits but low capacity).


```{r}
# Create a table with the percentage_tensed and the percentage_global across the four variables
tensed_summary <- tensed_hospitals %>%
  count(juridical_status, regulation, emergency_type, capacity) %>%
  group_by(juridical_status) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_status, by = "juridical_status") %>%
  group_by(regulation) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_reg, by = "regulation") %>%
  group_by(emergency_type) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_type, by = "emergency_type") %>%
  group_by(capacity) %>%
  mutate(percentage_tensed = n / sum(n) * 100) %>%
  left_join(global_cap, by = "capacity")

tensed_summary

# Save the tensed_summary table with gt
tensed_summary %>%
  gt() %>%
  tab_header(title = "Distribution of the Tensed Hospitals") %>%
  gtsave("../results/SAE_2024/tensed_hospitals_distribution.png")



```


# 3. Severity analysis: hosp_rate

We first want to know if those hossitals that have a high rate of patients per doctor and a low rate of doctors to population are also the ones with the highest hospitalization rates (which could be a proxy for severity of cases).

```{r}
# Scatterplot of hospitalization rate vs. patients per doctor by hospital (with quadrants based on the median)
ggplot(SAE_2024, aes(x = patients_per_doctor_win, y = hosp_rate_mod)) +
  geom_point(alpha = 0.6) +
  geom_vline(xintercept = median(SAE_2024$patients_per_doctor_win, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  geom_hline(yintercept = median(SAE_2024$hosp_rate_mod, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  labs(
    title = "Correlation between Patients per Doctor and Hospitalization Rate at the Hospital Level",
    subtitle = "Quadrants divided by median values",
    x = "Patients per Doctor (winsorized)",
    y = "Hospitalization Rate (modified)"
  ) +
  theme_minimal()
```

```{r}
# Check correlation
cor(tensed_hospitals$patients_per_doctor_win, tensed_hospitals$hosp_rate_mod, use = "complete.obs")
cor(SAE_2024$patients_per_doctor_win, SAE_2024$hosp_rate_mod, use = "complete.obs")

# Test correlation
cor.test(tensed_hospitals$patients_per_doctor_win, tensed_hospitals$hosp_rate_mod, use = "complete.obs")
cor.test(SAE_2024$patients_per_doctor_win, SAE_2024$hosp_rate_mod, use = "complete.obs")
```

Statistically significant negative correlation at the 1% threshold for both the whole dataset and the tensed hospitals dataset, ie the more patients per doctor, the lower the hospitalization rate.
The order of magnitude is different though: -0.15 for the whole dataset, -0.3 for the tensed hospitals dataset, which means that the correlation is stronger for the tensed hospitals than for the whole dataset.


We then study the characteristics of the hospitals in the top left corner of the scatterplot (low doctors per inhabitant, high patients per doctor) to see if they are also the ones with the highest hospitalization rates.

```{r}
# Scatterplot: x axis = doctors to population, y axis = patients per doctor, color (gradient) = hospitalization rate, with quadrants based on the median
hosp_scatter <- ggplot(SAE_2024, aes(x = dep_doctors_per_1000, y = patients_per_doctor_win, color = hosp_rate_mod)) +
  geom_point(alpha = 0.6) +
  geom_vline(xintercept = median(SAE_2024$dep_doctors_per_1000, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  geom_hline(yintercept = median(SAE_2024$patients_per_doctor_win, na.rm = TRUE), linetype = "dashed", color = "red", size = 0.8) +
  scale_color_viridis_c(option = "magma", name = "Hospitalization Rate (modified)") +
  labs(
    title = "Correlation between Doctors per Inhabitant, Patients per Doctor and Hospitalization Rate",
    subtitle = "Quadrants divided by median values",
    x = "Doctors per 1000 Inhabitants in the Department",
    y = "Patients per Doctor (winsorized)"
  ) +
  theme_minimal()

hosp_scatter
```

We isolate the hospitals in the top left corner (low doctors per inhabitant, high patients per doctor) and check if they are also the ones with the highest hospitalization rates.

```{r}
ggsave("../results/SAE_2024/hosp_scatter.png", hosp_scatter, width = 10, height = 8)
```

```{r}
# Zoom in the scatter plot in the top left corner
ggplot(tensed_hospitals, aes(x = dep_doctors_per_1000, y = patients_per_doctor_win, color = hosp_rate_mod)) +
  geom_point(alpha = 0.6) +
  scale_color_viridis_c(option = "magma", name = "Hospitalization Rate (modified)") +
  labs(
    title = "Hospitals with Low Doctors per Inhabitant and High Patients per Doctor",
    x = "Doctors per 1000 Inhabitants in the Department",
    y = "Patients per Doctor (winsorized)"
  ) +
  theme_minimal()
```

```{r}
# Identify hospitals from this scatterplot that have a high hosp rate (above the median of the tensed hospitals)
tensed_hospitals_2 <- tensed_hospitals %>%
  filter(hosp_rate_mod > median(tensed_hospitals$hosp_rate_mod, na.rm = TRUE)) %>%
  select(finess, establishment_name, dep_number, patients_per_doctor_win, hosp_rate_mod) %>%
  arrange(desc(hosp_rate_mod))

# Count the number of hospitals per department in the main dataset
hospitals_per_department <- SAE_2024 %>%
  count(dep_number) %>%
  rename(total_dep_hospitals = n)
```

We then plot those hospitals on a map to see where they are located in France.

```{r}
# 1. Calculate counts and rates per department
# We start by counting how many "tensed" hospitals are in each department
tensed_counts <- tensed_hospitals_2 %>%
  count(dep_number, name = "tensed_count")

# 2. Prepare the map data with the rate calculation
map_data <- departments_geo %>%
  # Join with the counts of tensed hospitals
  left_join(tensed_counts, by = c("code" = "dep_number")) %>%
  # Join with the total hospital counts per department (from your existing hospitals_per_department)
  left_join(hospitals_per_department, by = c("code" = "dep_number")) %>%
  # Replace NA with 0 for departments with no tensed hospitals
  mutate(tensed_count = replace_na(tensed_count, 0)) %>%
  # Calculate the rate
  mutate(tensed_rate = tensed_count / total_dep_hospitals) %>%
  # Filter out departments with no hospitals at all if necessary, or keep them to show 0%
  filter(!is.na(total_dep_hospitals)) %>% 
  # Ensure the point is inside the department
  mutate(centroid = st_point_on_surface(geometry),
         lon = st_coordinates(centroid)[,1],
         lat = st_coordinates(centroid)[,2])

map_data <-  map_data %>%
  arrange(desc(tensed_rate))

# 3. Create the Map using the rate
hosp_map <- ggplot() +
  geom_sf(data = departments_geo, fill = "gray95", color = "white") +
  geom_point(data = map_data, 
             aes(x = lon, y = lat, size = tensed_rate, color = tensed_rate),
             alpha = 0.8) +
  # Format the legend as percentages
  scale_size_continuous(name = "Tensed Hospital Rate", 
                        range = c(2, 12),
                        labels = scales::percent_format(accuracy = 1)) +
  scale_color_viridis_c(name = "Tensed Hospital Rate", 
                        option = "viridis",
                        labels = scales::percent_format(accuracy = 1)) +
  theme_void() +
  theme(legend.position = "right") +
  labs(title = "Share of Tensed Hospitals by Department",
       subtitle = "Percentage of hospitals with high hospitalization rate relative to department total")

hosp_map
```

```{r}
ggsave("../results/SAE_2024/hosp_map.png", hosp_map, width = 10, height = 8)
```

We study this rate across juridical status and regulation dimensions to see if the departments with the highest rate of tensed hospitals are also the ones with the highest share of private hospitals or the lowest regulation. TO DO


We also study the hospitalization rates as a proxy for the times when it was necessary to come to the emergency.

```{r}
map_data_general <- SAE_2024 %>%
  group_by(dep_number) %>%
  summarise(avg_hosp_rate = median(hosp_rate_mod, na.rm = TRUE), .groups = "drop") %>%
  left_join(departments_geo %>% select(code, geometry), by = c("dep_number" = "code")) %>%
  filter(!is.na(n)) %>%
  # Ensure the point is inside the department
  mutate(centroid = st_point_on_surface(geometry),
         lon = st_coordinates(centroid)[,1],
         lat = st_coordinates(centroid)[,2])

# Plot the map with points sized by hospitalization rate
ggplot() +
  geom_sf(data = departments_geo, fill = "gray", color = "white") +
  geom_point(data = map_data_general, 
             aes(x = lon, y = lat, color = avg_hosp_rate),
             alpha = 0.8) +
  #scale_size_continuous(name = "Hospitalization Rate (modified)", range = c(2, 12)) +
  scale_color_viridis_c(name = "Hospitalization Rate (modified)", option = "magma") +
  theme_void() +
  theme(legend.position = "right") +
  labs(title = "Hospitalization Rate (modified) by Hospital Location")

ggsave("../results/SAE_2024/hosp_rate_map.png", width = 10, height = 8)

```


# 4. Personnel analysis

```{r}

```




```{r}
# Import dataset with the total number of emergency doctors
Q20 <- read_csv2("../data/raw/SAE_2024/Q20_2024r.csv")

# Clean dataset
Q20 <-  Q20 %>%
  filter(PERSO == "M1340") %>%
  select(FI, AN, RS, EFFSAL) %>%
  rename(finess = FI)

# Merge with the main dataset
SAE_2024 <- SAE_2024 %>%
  left_join(Q20, by = "finess") %>%
  rename(total_number_doctors = EFFSAL) %>%
  select(-AN, -RS)

# Check NAs and values 0 in total_number_doctors
sum(is.na(SAE_2024$total_number_doctors))
sum(SAE_2024$total_number_doctors == 0, na.rm = TRUE)

# See if those are public
SAE_2024 %>%
  filter(is.na(total_number_doctors) | total_number_doctors == 0) %>%
  count(juridical_status) %>%
  mutate(percentage = n / sum(n) * 100)

# Checks rows where etp_physicians_empl_emergency is NA ANS total_number_doctors is 0 or NA
SAE_2024 %>%
  filter(is.na(etp_physicians_empl_emergency) & (is.na(total_number_doctors) | total_number_doctors == 0)) %>%
  count(juridical_status) %>%
  mutate(percentage = n / sum(n) * 100)
```





