---
title: "SAE_analysis"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

# Set up ----

```{r setup, message=FALSE}
rm(list=ls())

path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
library(ggplot2)
library(sf)
library(viridis)
library(scales)
```


# Import datasets ----

```{r cars, message=FALSE}
SAE_2024 <- read_csv2("../data/processed/SAE_final_2024.csv")
Legend <- read_csv2("../docs/SAE_2024/Legends_short.csv")
```

# Descriptive statistics: emergency services offer ----

## Are emergencies mostly public?

```{r, message = TRUE}
# Count the number of emergency services by status
status_counts <- SAE_2024 %>%
  group_by(juridical_status) %>%
  summarise(count = n())

status_counts
```

## Are there mostly CHU/CHRs or smaller hospitals?

```{r}
# Count the number of emergency services by size
size_counts <- SAE_2024 %>%
  group_by(size_es) %>%
  summarise(count = n())
```

## How are emergency services distributed geographically?

```{r}
# Count the number of emergency services by department
department_counts <- SAE_2024 %>%
  st_drop_geometry() %>%
  group_by(nom_dep) %>%
  summarise(count = n(), .groups = "drop")

summary(department_counts)
```


```{r}
# Aggregate the number of emergency services by department for mapping
map_data_es <- SAE_2024 %>%
  group_by(dep, nom_dep) %>%
  summarise(total_es = n(), .groups = "drop")

ggplot(data = map_data_es) +
  geom_sf(aes(fill = total_es), color = "white", size = 0.2) +
  scale_fill_viridis_c(
    option = "magma", 
    name = "Number of services",
    #breaks = c(1, 5, 10, 15, 20),
    #labels = c("1", "5", "10", "15", "20+")
  ) +
  labs(
    title = "Distribution of emergency services",
    subtitle = "Year 2024",
    caption = "Source : SAE – DREES"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```


# Descriptive statistics: medical activity ----
## How many passages in emergency services in a year?

```{r}
# Summary statistics for the number of passages in emergency services in a year
summary(SAE_2024$passu)
sd(SAE_2024$passu)
```

## How does the number of visits vary by establishment size?

```{r}
# Number of emergency visits by department
visits_count <- SAE_2024 %>%
  group_by(size_es) %>%
  summarise(total_passages = sum(passu, na.rm = TRUE))

# Summary statistics for the number of passages in emergency services in a year by establishment size
activity_by_size <- SAE_2024 %>%
  group_by(size_es) %>%
  summarise(
    n_hospitals = n(),
    mean_passages = mean(passu, na.rm = TRUE),
    sd_passages = sd(passu, na.rm = TRUE),
    median_passages = median(passu, na.rm = TRUE),
    min_passages = min(passu, na.rm = TRUE),
    max_passages = max(passu, na.rm = TRUE)
  )

activity_by_size

# Visualization of the number of passages by establishment size
ggplot(SAE_2024, aes(x = size_es, y = passu, fill = size_es)) +
  geom_boxplot() +
  labs(title = "Distribution du nombre de passages par an selon la taille de l'établissement",
       y = "Nombre de passages par an",
       x = "Taille de l'établissement") +
  theme_minimal()
```

## How does the number of passages vary by department?

```{r}
# Summary statistics for the number of passages in emergency services in a year by department
activity_by_department <- SAE_2024 %>%
  group_by(nom_dep) %>%
  summarise(
    n_hospitals = n(),
    mean_passages = mean(passu, na.rm = TRUE),
    sd_passages = sd(passu, na.rm = TRUE),
    median_passages = median(passu, na.rm = TRUE),
    min_passages = min(passu, na.rm = TRUE),
    max_passages = max(passu, na.rm = TRUE)
  )

# Number of emergency visits by department
visits_count <- SAE_2024 %>%
  st_drop_geometry() %>% 
  group_by(dep) %>%
  summarise(total_passages = sum(passu, na.rm = TRUE), .groups = "drop")
  
# Join with geo data for mapping
map_data_passages <- departments_geo %>%
  left_join(visits_count, by = c("code" = "dep"))

ggplot(data = map_data_passages) +
  geom_sf(aes(fill = total_passages), color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Number of visits (in a year)",
    labels = label_number(big.mark = " ", suffix = " pts")
  ) +
  labs(
    title = "Distribution of emergency service visits",
    subtitle = "Year 2024",
    caption = "Source : SAE – DREES"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```


# Descriptive statistics: tension indicators ----

# What is the number of patients per doctor (in a year)?

```{r}
# Convert passu to number/day
SAE_2024 <- SAE_2024 %>%
  mutate(
    passu_day = passu / 365
  )

# Create ratio variable
SAE_2024 <- SAE_2024 %>%
  mutate(
    patients_per_doctor = ifelse(etpsal_urg > 0, passu_day / etpsal_urg, NA)
  )

# Summary statistics for the number of patients per doctor in a year (exclude NAs/Infs)
summary(SAE_2024$patients_per_doctor[is.finite(SAE_2024$patients_per_doctor)])
#sd(SAE_2024$patients_per_doctor[is.finite(SAE_2024$patients_per_doctor)])

# Création du boxplot
ggplot(SAE_2024) +
  aes(x = size_es, y = patients_per_doctor, fill = size_es) +
  geom_boxplot(outlier.color = "red", outlier.shape = 1) +
  # On limite l'affichage à 100 pour y voir quelque chose, 
  # tout en gardant les données en mémoire
  coord_cartesian(ylim = c(0, 100)) + 
  labs(
    title = "Distribution de la charge de travail par médecin (par jour)",
    subtitle = "Zoom sur la zone 0-100 patients/jour (les points rouges sont des outliers)",
    x = "Taille de l'établissement",
    y = "Nombre de patients / jour / ETP médical",
    fill = "Catégorie"
  ) +
  theme_minimal()
```

# How does the number of patients per doctor vary by establishment size?

```{r}
# Summary statistics for the number of patients per doctor in a year by establishment size
activity_ratio_by_size <- SAE_2024 %>%
  group_by(size_es) %>%
  summarise(
    n_hospitals = n(),
    mean_ratio = mean(patients_per_doctor, na.rm = TRUE),
    sd_ratio = sd(patients_per_doctor, na.rm = TRUE),
    median_ratio = median(patients_per_doctor, na.rm = TRUE),
    min_ratio = min(patients_per_doctor, na.rm = TRUE),
    max_ratio = max(patients_per_doctor, na.rm = TRUE)
  )

activity_ratio_by_size

# Visualization of the number of patients per doctor by establishment size
ggplot(SAE_2024, aes(x = size_es, y = patients_per_doctor, fill = size_es)) +
  geom_boxplot() +
  # Limit Y values
  coord_cartesian(ylim = c(0, 10000)) +
  labs(title = "Distribution of the number of emergency visits per doctor, by establishment size",
       y = "Visits per doctor, in a year",
       x = "Establishment size") +
  theme_minimal()
```


# How does the number of patients per doctor vary by department?

```{r}
# Number of visits per doctor, by department
ratio_data <- SAE_2024 %>%
  st_drop_geometry() %>%
  group_by(dep) %>%
  summarise(
    total_passu = sum(passu, na.rm = TRUE),
    total_doctors = sum(etpsal_urg, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(patients_per_doctor = total_passu / total_doctors)

# Join with geo data for mapping
map_data_ratio <- departments_geo %>%
  left_join(ratio_data, by = c("code" = "dep"))

# Visualization of patients per doctor by department
ggplot(data = map_data_ratio) +
  geom_sf(aes(fill = patients_per_doctor), color = "white", size = 0.1) +
  scale_fill_viridis_c(
    option = "magma",
    name = "Visits / Doctors",
    na.value = "grey90",
    labels = label_number(big.mark = " ")
  ) +
  labs(
    title = "Distribution of the number of emergency visits per doctor, in a year",
    subtitle = "Year 2024",
    caption = "Source : SAE – DREES"
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )
```

# Bivariate analysis ----

```{r}
# Create table of the ratio of patients per doctor by establishment status (public, privé non lucratif, privé lucratif)
table_by_status <-  SAE_2024 %>%
  group_by(es_status) %>%
  summarise(
    nb_services = n(),
    mediane_passages = median(passu, na.rm = TRUE),
    mediane_ratio = median(patients_per_doctor, na.rm = TRUE),
    .groups = 'drop'
  )

# Visualization
ggplot(SAE_2024, aes(x = es_status, y = patients_per_doctor, fill = es_status)) +
  geom_boxplot() +
  # Limit Y values
  coord_cartesian(ylim = c(0, 10000)) + 
  labs(
    title = "Medical tension: public versus private emergengy services",
    subtitle = "Year 2024",
    x = "Juridical status",
    y = "Visits per doctor, in a year",
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```


```{r}
# Create table of the ratio of patients per doctor by emergency type (general, pediatric)
table_by_type <-  SAE_2024 %>%
  group_by(urg) %>%
  summarise(
    nb_services = n(),
    mediane_passages = median(passu, na.rm = TRUE),
    mediane_ratio = median(patients_per_doctor, na.rm = TRUE),
    .groups = 'drop'
  )

# Visualization
ggplot(SAE_2024, aes(x = urg, y = patients_per_doctor, fill = urg)) +
  geom_boxplot() +
  # Limit Y values
  coord_cartesian(ylim = c(0, 10000)) + 
  labs(
    title = "Medical tension: general versus pediatric emergency services",
    subtitle = "Year 2024",
    x = "Type of emergency service",
    y = "Visits per doctor, in a year",
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```