---
title: "SAE_cleaning_NA"
author: "Louise Gatty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
path = "/Users/Louise/Desktop/Master/M1/PiA-DREES-local/code"
setwd(path)

# Load necessary libraries ----
library(readxl)
library(dplyr)
library(tidyr)
library(readr)
library(httr)
library(readr)
library(janitor)
library(tidyverse)
library(stringr)
library(purrr)
library(ggplot2)
library(sf)
library(viridis)
library(scales)
library(tibble)
```

# Import datasets ----

```{r cars, message=FALSE}
SAE_2024 <- read_csv2("../data/intermediate/SAE_merged_short_2024.csv")
Legend <- read_csv2("../docs/SAE_2024/Legends_short.csv")
```

# Data cleaning

## Redefine juridical status

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create a new variable for juridical status
  mutate(juridical_status = case_match(stjr,
    1 ~ "Public",
    2 ~ "Privé non lucratif",
    3 ~ "Privé lucratif",
  )) %>%
  # Drop the initial juridical status variable
  select(-stjr)
```

## Redefine establishment capacity

Catr gives more information on the juridical status than on the establishment size. We prefer using the number of beds as an indicator of the size.

We can use 2 figures: the number of beds in the emergency service and the number of beds in the whole establishment.

Below, we try to use the number of beds in the whole establishment. The reasoning for it is that a lot of emergency services suffer from a lack of beds on the wards (medical, surgical) to transfer patients.

```{r}
# Import data that contains the overall number of beds
usld <- read_csv2("../data/raw/SAE_2024/USLD_2024r.csv")

# Create temporary dataframe of the joined data of SAE and usld
temp <-  SAE_2024 %>%
  left_join(usld, by = c("fi" = "FI"))

# Compute percentage of NA values
sum(is.na(temp$LIT)) / length(temp$LIT) * 100

# Define function to group the NAs
analyze_missing_pct <- function(data, group_var, target_var) {
  data %>%
    group_by({{ group_var }}) %>%
    summarise(
      missing_percentage = sum(is.na({{ target_var }})) / n() * 100
    ) %>%
    arrange(desc(missing_percentage))
}

# Apply to LIT
analyze_missing_pct(temp, juridical_status, LIT)
```

The percentage of missing values for the overall number of beds is very high (74%), even for public hospitals (67%).

The emergency beds number is rather an indicator of capacity of the emergency service. It does not exactly answer the same question, but it will allow us to identify the establishments for which the capacity is small but the flow is high.

```{r}
rm(usld)
rm(temp)

# Compute thresholds
terciles <- quantile(SAE_2024$lit_uhcd, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)

# Create categorical variable
SAE_2024 <- SAE_2024 %>%
  mutate(capacity = case_when(
    lit_uhcd <= terciles[2] ~ "Low capacity",
    lit_uhcd <= terciles[3] ~ "Medium capacity",
    lit_uhcd >  terciles[3] ~ "High capacity"
  )) %>%
  # Lock the order
  mutate(capacity = factor(capacity, 
                                    levels = c("Low capacity", "Medium capacity", "High capacity")))

# View establishments with NAs for the number of beds
temp <- SAE_2024 %>%
  #filter(is.na(lit_uhcd)) %>%
  select(rs, catr, capacity, dep, passu, lit_uhcd)
```

Deleting these observations, even if there is a small number of them, would false (eg.: Vendée would loose its biggest hospital).
We use the catr variable to impute the average number of beds of the same categories' establishments in the missing values.

```{r}
rm(temp)

# Mean by catr
mean_by_catr <- SAE_2024 %>%
  group_by(catr) %>%
  summarise(mean_lit_uhcd_catr = mean(lit_uhcd, na.rm = TRUE), .groups = "drop")

# Global mean in case one of the catr has a 100% of NAs
global_mean <- mean(SAE_2024$lit_uhcd, na.rm = TRUE)

# Imputation
SAE_2024 <- SAE_2024 %>%
  left_join(mean_by_catr, by = "catr") %>%
  mutate(
    lit_uhcd_imp = if_else(
      is.na(lit_uhcd),
      # si mean du catr est NaN (groupe sans valeurs), on met la moyenne globale
      if_else(is.nan(mean_lit_uhcd_catr), global_mean, mean_lit_uhcd_catr),
      lit_uhcd
    )
  ) %>%
  select(-mean_lit_uhcd_catr)

# Recompute the terciles
terciles_imp <- quantile(SAE_2024$lit_uhcd_imp, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE)

# Create categorical variable
SAE_2024 <- SAE_2024 %>%
  mutate(
    capacity = case_when(
      lit_uhcd_imp <= terciles_imp[2] ~ "Low capacity",
      lit_uhcd_imp <= terciles_imp[3] ~ "Medium capacity",
      lit_uhcd_imp >  terciles_imp[3] ~ "High capacity"
    ),
    # Lock in the order
    capacity = factor(capacity,
                      levels = c("Low capacity", "Medium capacity", "High capacity"))
  ) %>%
  # Drop lit_uhcd and catr
  select(-lit_uhcd) %>%
  select(-catr)

# Remove temporary datasets
rm(mean_by_catr)
rm(terciles)
rm(terciles_imp)
```

## Join with department and region geographical data

```{r}
# Import geojson files
departments_geo <- st_read("../data/raw/departements.geojson")
regions_geo <- st_read("../data/raw/regions.geojson")

# Prepare region file for merging
regions_to_join <- regions_geo %>% 
  rename(region_geometry = geometry) %>% 
  rename(region_name = nom) %>%
  as.data.frame()

# Prepare department file for merging
departments_geo <- departments_geo %>%
  rename(dep_geometry = geometry) %>%
  rename(dep_name = nom)

# Merge with region and department files
SAE_2024 <- SAE_2024 %>%
  left_join(departments_geo, by = c("dep" = "code")) %>%
  left_join(regions_to_join, by = c("reg" = "code")) %>%
  st_as_sf()
```

## Rename variables for clearer understanding

```{r}
# Rename the variables in SAE_2024
SAE_2024 <- SAE_2024 %>%
  rename(year = an) %>%
  rename(finess = fi) %>%
  rename(establishment_name = rs) %>%
  rename(dep_number = dep) %>%
  rename(region_number = reg) %>%
  rename(town_name = nomcom) %>%
  rename(postal_code = cpo) %>%
  rename(emergency_type = urg) %>%
  rename(total_emergency_visits = passu) %>% # total for a year
  rename(total_hospitalized_emergency = dt_hosp) %>% # total for a year
  rename(waiting_monitoring = stap) %>%
  rename(beds_number = lit_uhcd_imp) %>% # installed as of December 31 in short-stay hospitalization units (UHCD)
  rename(hours_physicians_total = hmed_tot) %>% # weekly average, Urgences + SMUR + SAMU
  rename(hours_physicians_emergency = hmed_urg) %>% # weekly average, Urgences only
  rename(hours_nurses_total = hide_tot) %>% # weekly average, Urgences + SMUR + SAMU
  rename(hours_nurses_emergency = hide_urg) %>% # weekly average, Urgences only
  rename(hours_paramedics_total = hamb_tot) %>% # weekly average, Urgences + SMUR + SAMU
  rename(hours_regulation_physicians = hmr_tot) %>% # weekly average, Urgences + SMUR + SAMU
  rename(hours_regulation_amu_phyisicians = hmramu_tot) %>% # weekly average, physicians participating in emergency medical assistance (aide médical d'urgence), Urgences + SMUR + SAMU
  rename(hours_regulation_assistants = arm_tot) %>% # weekly average, Urgences + SMUR + SAMU
  rename(hours_onsp = osnp_tot) %>% # weekly average, Non-scheduled healthcare operators (OSNP), Urgences + SMUR + SAMU
  rename(joint_personnel_SMUR_emergency = com_smur_su) %>%
  rename(etp_physicians_empl_total = etpsal_tot) %>%
  rename(etp_physicians_empl_emergency = etpsal_urg) %>%
  rename(number_physicians_liberal_total = efflib_tot) %>%
  rename(number_physicians_liberal_emergency = efflib_urg) %>%
  rename(number_medecine_interns = intmed) %>%
  rename(number_emergency_interns = inturg) %>%
  rename(number_medecine_junior_doctors = docju_med) %>%
  rename(number_emergency_junior_doctors = docju_urg)
  
# Rearrange the order
SAE_2024 <- SAE_2024 %>%
  select(year, finess, establishment_name, juridical_status, capacity, region_number, region_name, region_geometry, dep_number, dep_name, dep_geometry, everything())

# Remove the unnecessary datasets for the further cleaning
rm(departments_geo, regions_geo, regions_to_join)
```

Overseas territories were not present in the geojson department file => we rewrite the names and geometry of the region in the department names, for departments 9A 9B 9C 9D 9F.

```{r}
# Identify overseas territories lines
dom_indices <- which(str_detect(SAE_2024$dep_number, "^9[A-F]"))

# Replace department name (NAs) by region name
SAE_2024$dep_name[dom_indices] <- SAE_2024$region_name[dom_indices]

# Idem for the geometry
st_geometry(SAE_2024)[dom_indices] <- st_geometry(SAE_2024$region_geometry)[dom_indices]

# Ensure that SAE_2024 is recognize as spatial 
SAE_2024 <- SAE_2024 %>%
  st_as_sf()

rm(dom_indices)
```

## NAs

```{r}
# Count the percentage of missing values for each variable in SAE_2024
missing_percentage <- colSums(is.na(SAE_2024)) / nrow(SAE_2024) * 100
missing_percentage
```

### Explore patterns of NA observations

#### Hours_regulation_physicians

```{r}
sum(is.na(SAE_2024$hours_regulation_physicians)) / length(SAE_2024$hours_regulation_physicians) * 100

analyze_missing_pct(SAE_2024, capacity, hours_regulation_physicians)
analyze_missing_pct(SAE_2024, juridical_status, hours_regulation_physicians)
analyze_missing_pct(SAE_2024, region_name, hours_regulation_physicians)
```
Very high level of missing values (79%). A little less NAs in public/high capacity establishments. Missing values percentage of at least 50 % in 17 regions out of 18.
Since the variable seems core for the analysis, we create a dummy for it.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_regulation_physicians = if_else(is.na(hours_regulation_physicians), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(hours_regulation_physicians = replace_na(hours_regulation_physicians, 0))
```

#### Hours_paramedics_total

```{r}
sum(is.na(SAE_2024$hours_paramedics_total)) / length(SAE_2024$hours_paramedics_total) * 100

analyze_missing_pct(SAE_2024, capacity, hours_paramedics_total)
analyze_missing_pct(SAE_2024, juridical_status, hours_paramedics_total)
```

41 % of NAs overall.
Seems to be missing for small/medium capacity and private establishments mainly, which makes sense => create dummy.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_paramedics = if_else(is.na(hours_paramedics_total), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(hours_paramedics_total = replace_na(hours_paramedics_total, 0))
```

#### Hours_regulation_amu_phyisicians

```{r}
sum(is.na(SAE_2024$hours_regulation_amu_phyisicians)) / length(SAE_2024$hours_regulation_amu_phyisicians) * 100

analyze_missing_pct(SAE_2024, capacity, hours_regulation_amu_phyisicians)
analyze_missing_pct(SAE_2024, juridical_status, hours_regulation_amu_phyisicians)
analyze_missing_pct(SAE_2024, region_name, hours_paramedics_total)
```

81% of NAs overall. No clear pattern (highly rate of missing values for public and high capacity establishments too, contrary to the previous variable).
Variable does not seem essential => drop it.

```{r}
SAE_2024 <-  SAE_2024 %>%
  select(-hours_regulation_amu_phyisicians)
```


#### Hours_regulation_assistants

```{r}
sum(is.na(SAE_2024$hours_regulation_assistants)) / length(SAE_2024$hours_regulation_assistants) * 100

analyze_missing_pct(SAE_2024, capacity, hours_regulation_assistants)
analyze_missing_pct(SAE_2024, juridical_status, hours_regulation_assistants)
analyze_missing_pct(SAE_2024, region_name, hours_regulation_assistants)
```

79% of NAs overall. No very clear pattern. However the variable seems essential for the analysis => dummy.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_regulation_assistants = if_else(is.na(hours_regulation_assistants), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(hours_regulation_assistants = replace_na(hours_regulation_assistants, 0))

# Check for perfect colinarity with dummy_regulation_physicians
cor(SAE_2024$dummy_regulation_physicians, SAE_2024$dummy_regulation_assistants)
```

0.96 correlation => highly correlated but not problematic for regression analysis.

#### Hours_onsp

```{r}
sum(is.na(SAE_2024$hours_onsp)) / length(SAE_2024$hours_onsp) * 100

analyze_missing_pct(SAE_2024, capacity, hours_onsp)
analyze_missing_pct(SAE_2024, juridical_status, hours_onsp)
analyze_missing_pct(SAE_2024, region_name, hours_onsp)
```
86% of NAs overall. No clear pattern. Variable seems less core for the analysis => drop it.

```{r}
SAE_2024 <-  SAE_2024 %>%
  select(-hours_onsp)
```


#### Number_physicians_liberal_total and number_physicians_liberal_emergency

```{r}
# Number_physicians_liberal_total
sum(is.na(SAE_2024$number_physicians_liberal_total)) / length(SAE_2024$number_physicians_liberal_total) * 100

analyze_missing_pct(SAE_2024, capacity, number_physicians_liberal_total)
analyze_missing_pct(SAE_2024, juridical_status, number_physicians_liberal_total)
analyze_missing_pct(SAE_2024, region_name, number_physicians_liberal_total)
```
74% of NAs overall. 
Very high level of NAs except for non lucrative private establishments. 100% missing for most overseas territories.
The variable remains interesting since it's correlated to the juridical status => dummy.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_number_liberal_phyisicians = if_else(is.na(number_physicians_liberal_total), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_physicians_liberal_total = replace_na(number_physicians_liberal_total, 0))
```

```{r}
# Number_physicians_liberal_emergency
sum(is.na(SAE_2024$number_physicians_liberal_emergency)) / length(SAE_2024$number_physicians_liberal_emergency) * 100

analyze_missing_pct(SAE_2024, capacity, number_physicians_liberal_emergency)
analyze_missing_pct(SAE_2024, juridical_status, number_physicians_liberal_emergency)
analyze_missing_pct(SAE_2024, region_name, number_physicians_liberal_emergency)
```
75% of NAs overall. 
Very high level of NAs except non lucrative private establishments. 100% missing for most overseas territories.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_number_liberal_phyisicians_emergency = if_else(is.na(number_physicians_liberal_emergency), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_physicians_liberal_emergency = replace_na(number_physicians_liberal_emergency, 0))

# Check for correlation with dummy_number_liberal_physicians
cor(SAE_2024$dummy_number_liberal_phyisicians_emergency, SAE_2024$dummy_number_liberal_phyisicians)
```

0.98 correlation. Urgences only and Urgences+SMUR+SAMU are quasi similar for small establishments (no SMUR nor SAMU) => less information given by number_physicians_liberal_emergency in addition to dummy_number_liberal_physicians (and the other way around).

=> drop number_physicians_liberal_emergency (since its percentage of NAs is 1 pp higher) and the associated dummy.

```{r}
SAE_2024 <-  SAE_2024 %>%
  select(-c(dummy_number_liberal_phyisicians_emergency, number_physicians_liberal_emergency))
```

#### Joint_personnel_SMUR_emergency

```{r}
sum(is.na(SAE_2024$joint_personnel_SMUR_emergency)) / length(SAE_2024$joint_personnel_SMUR_emergency) * 100

analyze_missing_pct(SAE_2024, capacity, joint_personnel_SMUR_emergency)
analyze_missing_pct(SAE_2024, juridical_status, joint_personnel_SMUR_emergency)
```

37% of NAs overall.
NAs mainly in private establishments, for which there's a high chance this question does not apply.
Variable is interesting as a measurement of the workload of emergency personnel => dummy.

```{r}
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_joint_personnel = if_else(is.na(joint_personnel_SMUR_emergency), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(joint_personnel_SMUR_emergency = replace_na(joint_personnel_SMUR_emergency, 0))
```

#### Interns

```{r}
# Medicine interns
sum(is.na(SAE_2024$number_medecine_interns)) / length(SAE_2024$number_medecine_interns) * 100

analyze_missing_pct(SAE_2024, capacity, number_medecine_interns)
analyze_missing_pct(SAE_2024, juridical_status, number_medecine_interns)

# Emergency interns
sum(is.na(SAE_2024$number_emergency_interns)) / length(SAE_2024$number_emergency_interns) * 100

analyze_missing_pct(SAE_2024, capacity, number_emergency_interns)
analyze_missing_pct(SAE_2024, juridical_status, number_emergency_interns)
```

Smaller percentage of missing data for public and high capacity establishments. Missing data is likely to be due to the absence of interns in the establishment => dummy.

```{r}
# Medecine interns
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_medecine_interns = if_else(is.na(number_medecine_interns), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_medecine_interns = replace_na(number_medecine_interns, 0))

# Emergency interns
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_emergency_interns = if_else(is.na(number_emergency_interns), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_emergency_interns = replace_na(number_emergency_interns, 0))

# Check correlation between the dummies
cor(SAE_2024$dummy_emergency_interns, SAE_2024$dummy_medecine_interns)
```

0.59 correlation. We keep both dummies.

#### Junior doctors

```{r}
# Medicine junior doctors
sum(is.na(SAE_2024$number_medecine_junior_doctors)) / length(SAE_2024$number_medecine_junior_doctors) * 100

analyze_missing_pct(SAE_2024, capacity, number_medecine_junior_doctors)
analyze_missing_pct(SAE_2024, juridical_status, number_medecine_junior_doctors)

# Emergency junior doctors
sum(is.na(SAE_2024$number_emergency_junior_doctors)) / length(SAE_2024$number_emergency_junior_doctors) * 100

analyze_missing_pct(SAE_2024, capacity, number_emergency_junior_doctors)
analyze_missing_pct(SAE_2024, juridical_status, number_emergency_junior_doctors)
```

51% and 71% missing values. Low (medicine) or lower (emergency) percentage for public establishments => dummy.

```{r}
# Medecine junior doctors
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_medecine_junior_doctors = if_else(is.na(number_medecine_junior_doctors), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_medecine_junior_doctors = replace_na(number_medecine_junior_doctors, 0))

# Emergency junior doctors
SAE_2024 <- SAE_2024 %>%
  # Create dummy
  mutate(dummy_emergency_junior_doctors = if_else(is.na(number_emergency_junior_doctors), 0, 1)) %>%
  # Replace the NAs by zero in the original variable
  mutate(number_emergency_junior_doctors = replace_na(number_emergency_junior_doctors, 0))

# Check correlation between the dummies
cor(SAE_2024$dummy_medecine_junior_doctors, SAE_2024$dummy_emergency_junior_doctors)
```

0.56 correlation. We keep both dummies.

# Save cleaned dataset ----

```{r}
# Drop geometrical variables for export
SAE_2024 <-  SAE_2024 %>%
  select(-region_geometry) %>%
  select(-dep_geometry)

# Export to csv
SAE_2024 %>%
  st_drop_geometry() %>% 
  write_csv2("../data/intermediate/SAE_cleaned_NA_2024.csv")
```